---
title: "PEP725 explorations"
author: "Beni Stocker"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(rbeni)
```

## Explorations

Just one file.
```{r eval=FALSE}
df <- read_delim("~/data/pep/PEP725_2020-12/pep725_2020-12_10000000.csv", delim = ";")
df_site <- read_delim("~/data/pep/PEP725_2020-12/pep725_datasource_2020-12_10000000.csv", delim = ";")
```

```{r eval=FALSE}
skimr::skim(df)
print(df)
```

## Reproducing Zani

Read all files.
```{r}
vec_files <- list.files("~/data/pep/PEP725_2020-12/", pattern = "pep725_2020-12_", full.names = TRUE)
list_df <- map(as.list(vec_files), ~read.delim(., sep = ";"))
names(list_df) <- vec_files

df <- list_df %>% 
  bind_rows(.id = "filnam") %>% 
  as_tibble() %>%  
  mutate(date = ymd(date))
```


### Interpretation of phenological states

Zani et al.:

*Phenology definitions followed the BBCH (Biologische Bundesanstalt, Bundessortenamt und Chemische Industrie) codes (see table S1). For the five deciduous angiosperms, spring leaf-out was defined as the date when the first or 50% of leaf stalks are visible (BBCH11 or BBCH13, respectively), for Larix decidua leaf-out was defined as the date when the first leaves separated (mouse ear stage = BBCH10). Leaf senescence was defined as the date when 50% of leaves had lost their green color (BBCH94) or had fallen (BCCH95).*

BBCH94 is not available in this dataset. Using BCCH205 instead (autumnal colouration >=50%)

This information is given in column `phase_id`.
```{r eval=FALSE}
df_sub <- df %>% 
  filter(phase_id %in% c(10, 11, 13, 95, 205)) %>% 
  mutate(id = paste0("i", 1:n())) %>% 
  mutate(pheno = ifelse(phase_id %in% c(10, 11, 13), "on", ifelse(phase_id %in% c(95, 205), "off", NA))) %>% 
  mutate(date = ymd(date)) %>% 
  mutate(lon = round(lon, digits = 4), lat = round(lat, digits = 4)) %>% 
  na_if(-9999) %>% 
  mutate(elv = ifelse(is.na(alt), alt_dem, alt))
```

Determine unique "sites".
```{r eval=FALSE}
df_sites <- df_sub %>% 
  select(lon, lat, elv) %>% 
  distinct() %>% 
  mutate(id_site = paste0("site_i", 1:n()))

df_sub <- df_sub %>% 
  left_join(df_sites, by = c("lon", "lat", "elv"))
```

Determine unique species x "sites".
```{r eval=FALSE}
df_species_sites <- df_sub %>% 
  select(lon, lat, elv, species) %>% 
  distinct() %>% 
  mutate(id_species_site = paste0("species_site_i", 1:n()))

df_sub <- df_sub %>% 
  left_join(df_species_sites, by = c("lon", "lat", "elv", "species"))
```

Determine unique species x "sites" x pheno-event.
```{r eval=FALSE}
df_pheno_species_sites <- df_sub %>% 
  select(lon, lat, elv, species, pheno) %>% 
  distinct() %>% 
  mutate(id_pheno_species_site = paste0("pheno_species_site_i", 1:n()))

df_sub <- df_sub %>% 
  left_join(df_pheno_species_sites, by = c("lon", "lat", "elv", "species", "pheno"))
```


### Cleaning

Remove individual time series with less than 15 years of leaf-out and leaf senescence observations,
```{r eval=FALSE}
n_years <- function(vec){length(unique(vec))}
df_retain <- df_sub %>% 
  group_by(id_species_site) %>% 
  summarise(n_years = n_years(year)) %>% 
  filter(n_years >= 15)

df_sub <- df_sub %>% 
  filter(id_species_site %in% unique(df_retain$id_species_site))
```

Remove dates deviating from an individual’s median more than 3 times the median absolute deviation.
```{r eval=FALSE}
medabsdev <- function(vec){abs(vec - median(vec)) / median(abs(vec - median(vec)))}

df_retain <- df_sub %>% 
  group_by(id_pheno_species_site, pheno) %>% 
  summarise(medabsdev = medabsdev(day)) %>% 
  filter(medabsdev < 3)

df_sub <- df_sub %>% 
  filter(id_pheno_species_site %in% unique(df_retain$id_pheno_species_site))
```

Remove time series for which the standard deviation of phenological observations across years was higher than 15 for leaf-out and 20 for leaf senescence.
```{r eval=FALSE}
df_retain <- df_sub %>% 
  group_by(id_pheno_species_site, pheno) %>% 
  summarise(sd = sd(day)) %>% 
  filter((sd <= 15.0 & pheno == "on") | (sd <= 20.0 & pheno == "off"))

df_sub <- df_sub %>% 
  filter(id_pheno_species_site %in% unique(df_retain$id_pheno_species_site))
```

Save intermediate file.
```{r eval=FALSE}
write_csv(df_sub, path = "data/df_sub.RData")
```

### Analysis

Read if not yet available.
```{r}
df_sub <- read_csv("data/df_sub.RData")
```

```{r eval=FALSE}
## this creates duplicates - problem!!!
df_anl <- df_sub %>% 
  select(id_species_site, pheno, day, year) %>% 
  pivot_wider(names_from = "pheno", values_from = "day")

## solution, but reduces from 2.6 to 2.2 mio obs.
df_anl <- df_sub %>% 
  distinct(id_pheno_species_site, year, .keep_all = TRUE) %>% 
  select(id_site, id_species_site, pheno, day, year) %>% 
  pivot_wider(names_from = "pheno", values_from = "day") %>% 
  filter(!is.na(on) & !is.na(off))
```

## Interannual

Get correlation between on and off for withing species-sites => interannual temporal variation.
```{r eval=FALSE}
get_coef_on <- function(df){df %>% filter(term == "on") %>% pull(estimate)}
get_p_on <- function(df){df %>% filter(term == "on") %>% pull(p.value)}

df_temporal <- df_anl %>% 
  # ungroup() %>% 
  # slice(1:10) %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_temporal, file = "data/df_temporal.RData")
```

```{r}
df_temporal %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red")
```

## Long-term

Get correlation between on and off for withing species-sites => long-term temporal variation (30 bins).
```{r eval=FALSE}
df_temporal_12 <- df_anl %>% 
  mutate(yearbin = ntile(year, n = 12)) %>% 
  group_by(id_species_site, yearbin) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_temporal_12, file = "data/df_temporal_12.RData")
```

```{r}
df_temporal_12 %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red")
```

## Spatial

Get correlation between on and off for withing species-sites => long-term temporal variation (30 bins).
```{r eval=FALSE}
df_spatial <- df_anl %>% 
  group_by(id_site) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE))

save(df_spatial, file = "data/df_spatial.RData")
```

```{r}
out_modobs <- df_spatial %>% analyse_modobs2("on", "off", type = "heat")
out_modobs$gg
```

## Spatial by species

Get correlation between on and off for withing species-sites => long-term temporal variation (30 bins).
```{r eval=FALSE}
df_spatial_species <- df_anl %>% 
  left_join(select(df_sub, id_species_site, species), by = c("id_species_site")) %>% 
  group_by(id_site, species) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>% 
  
  ungroup() %>% 
  group_by(species) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_spatial_species, file = "data/df_spatial_species.RData")
```

```{r}
df_spatial_species %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram(binwidth = 0.5) +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_spatial_species$coef_on, na.rm = TRUE), color = "springgreen3")
```

## Summary

```{r}
df_temporal %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red") +
  geom_vline(xintercept = median(df_spatial_species$coef_on, na.rm = TRUE), color = "springgreen3") +
  geom_vline(xintercept = out_modobs$df_metrics %>% filter(.metric == "slope") %>% pull(.estimate), color = "royalblue")
```