---
title: "PEP725 explorations"
author: "Beni Stocker"
date: "12/2/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(rbeni)
library(lme4) #lmer
library(MuMIn) #r.squaredGLMM
library(lmerTest) #p-values
library(effects) #plot effects
library(sjPlot) #plot effects
do_eval <- FALSE
```

## Explorations

This is to investigate main points using own analysis and code.

Read all files.
```{r eval=do_eval}
vec_files <- list.files("~/data/pep/PEP725_2020-12/", pattern = "pep725_2020-12_", full.names = TRUE)
list_df <- map(as.list(vec_files), ~read.delim(., sep = ";"))
names(list_df) <- vec_files

df <- list_df %>% 
  bind_rows(.id = "filnam") %>% 
  as_tibble() %>%  
  mutate(date = ymd(date))
```

### Interpretation of phenological states

Zani et al.:

*Phenology definitions followed the BBCH (Biologische Bundesanstalt, Bundessortenamt und Chemische Industrie) codes (see table S1). For the five deciduous angiosperms, spring leaf-out was defined as the date when the first or 50% of leaf stalks are visible (BBCH11 or BBCH13, respectively), for Larix decidua leaf-out was defined as the date when the first leaves separated (mouse ear stage = BBCH10). Leaf senescence was defined as the date when 50% of leaves had lost their green color (BBCH94) or had fallen (BCCH95).*

BBCH94 is not available in this dataset. Using BCCH205 instead (autumnal colouration >=50%)

This information is given in column `phase_id`.
```{r eval=do_eval}
df_sub <- df %>% 
  filter(phase_id %in% c(10, 11, 13, 95, 205)) %>% 
  mutate(id = paste0("i", 1:n())) %>% 
  mutate(pheno = ifelse(phase_id %in% c(10, 11, 13), "on", ifelse(phase_id %in% c(95, 205), "off", NA))) %>% 
  mutate(date = ymd(date)) %>% 
  mutate(lon = round(lon, digits = 4), lat = round(lat, digits = 4)) %>% 
  na_if(-9999) %>% 
  mutate(elv = ifelse(is.na(alt), alt_dem, alt))
```

Determine unique "sites".
```{r eval=do_eval}
df_sites <- df_sub %>% 
  select(lon, lat, elv) %>% 
  distinct() %>% 
  mutate(id_site = paste0("site_i", 1:n()))

df_sub <- df_sub %>% 
  left_join(df_sites, by = c("lon", "lat", "elv"))
```

Determine unique species x "sites".
```{r eval=do_eval}
df_species_sites <- df_sub %>% 
  select(lon, lat, elv, species) %>% 
  distinct() %>% 
  mutate(id_species_site = paste0("species_site_i", 1:n()))

df_sub <- df_sub %>% 
  left_join(df_species_sites, by = c("lon", "lat", "elv", "species"))
```

Determine unique species x "sites" x pheno-event.
```{r eval=do_eval}
df_pheno_species_sites <- df_sub %>% 
  select(lon, lat, elv, species, pheno) %>% 
  distinct() %>% 
  mutate(id_pheno_species_site = paste0("pheno_species_site_i", 1:n()))

df_sub <- df_sub %>% 
  left_join(df_pheno_species_sites, by = c("lon", "lat", "elv", "species", "pheno"))
```

### Cleaning

Remove individual time series with less than 15 years of leaf-out and leaf senescence observations,
```{r eval=do_eval}
n_years <- function(vec){length(unique(vec))}
df_retain <- df_sub %>% 
  group_by(id_species_site) %>% 
  summarise(n_years = n_years(year)) %>% 
  filter(n_years >= 15)

df_sub <- df_sub %>% 
  filter(id_species_site %in% unique(df_retain$id_species_site))
```

Remove dates deviating from an individual’s median more than 3 times the median absolute deviation.
```{r eval=do_eval}
medabsdev <- function(vec){abs(vec - median(vec)) / median(abs(vec - median(vec)))}

df_retain <- df_sub %>% 
  group_by(id_pheno_species_site, pheno) %>% 
  summarise(medabsdev = medabsdev(day)) %>% 
  filter(medabsdev < 3)

df_sub <- df_sub %>% 
  filter(id_pheno_species_site %in% unique(df_retain$id_pheno_species_site))
```

Remove time series for which the standard deviation of phenological observations across years was higher than 15 for leaf-out and 20 for leaf senescence.
```{r eval=do_eval}
df_retain <- df_sub %>% 
  group_by(id_pheno_species_site, pheno) %>% 
  summarise(sd = sd(day)) %>% 
  filter((sd <= 15.0 & pheno == "on") | (sd <= 20.0 & pheno == "off"))

df_sub <- df_sub %>% 
  filter(id_pheno_species_site %in% unique(df_retain$id_pheno_species_site))
```

Save intermediate file.
```{r eval=do_eval}
write_csv(df_sub, path = "~/data/df_sub.csv")
```

### Analysis

Read if not yet available. 

This is based on my own implementation of the data cleaning:
```{r}
df_sub <- read_csv("~/data/df_sub.csv")
```

This is based on the Zani et al. implementation (run by `zani.Rmd`):
```{r}
df_sub <- data.table::fread("data/DataMeta_2_PhenologyObs_PEP725_CleanData.csv") %>% 
  as_tibble() %>% 
  mutate(pheno = ifelse(phenology == "leaf.out", "on", ifelse(phenology == "leaf.off", "off", NA))) %>% 
  mutate(id_species_site = timeseries) %>% 
  rename(day = DoY, id_site = s_id)
```

```{r eval=do_eval}
## this creates duplicates - problem!!!
df_anl <- df_sub %>% 
  select(id_species_site, pheno, day, year) %>% 
  pivot_wider(names_from = "pheno", values_from = "day")

## solution, but reduces from 2.6 to 2.2 mio obs.
df_anl <- df_sub %>% 
  distinct(id_pheno_species_site, year, .keep_all = TRUE) %>% 
  select(id_site, id_species_site, pheno, day, year) %>% 
  pivot_wider(names_from = "pheno", values_from = "day") %>% 
  filter(!is.na(on) & !is.na(off))

write_csv(df_anl, path = "data/df_anl_zani.csv")
```

**Laura:** For the mixed effects modelling, you may best work with `df_anl` (I think).

Coverage of years in the data.
```{r}
df_anl <- read_csv("~/data/df_anl_zani.csv")
df_anl <- df_anl %>% mutate(species=stringr::str_split(id_species_site, "_") %>% map_chr(., 2))
df_anl %>% 
  ggplot(aes(year, ..count..)) +
  geom_histogram() +
  labs(title = "Number of data points per year", subtitle = "By sites and species, providing both EOS and SOS")
```

## Interannual

Get correlation between on and off for withing species-sites => interannual temporal variation.
```{r eval=do_eval}
df_anl <- read_csv("data/df_anl.csv")
get_coef_on <- function(df){df %>% filter(term == "on") %>% pull(estimate)}
get_p_on <- function(df){df %>% filter(term == "on") %>% pull(p.value)}

df_temporal <- df_anl %>% 
  ## for each species and site
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_temporal, file = "~/data/df_temporal.RData")
```

Interannual temporal variation with LMMs
```{r}
length(unique(df_anl$id_species_site))
length(unique(df_anl$id_site))
length(unique(df_anl$species))
length(unique(df_anl$year))

# Both models are equivalent
Fit_IAV1 = lmer(off ~ on + (1|id_site/species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV1)
r.squaredGLMM(Fit_IAV1)
plot(allEffects(Fit_IAV1))
Fit_IAV2 = lmer(off ~ on + (1|id_species_site) + (1|id_site), data = df_anl, na.action = "na.exclude")
summary(Fit_IAV2)
r.squaredGLMM(Fit_IAV2)
plot(allEffects(Fit_IAV2))

# Analysis of residuals
ResG <- residuals(Fit_IAV1)
FittedG <- fitted(Fit_IAV1)
par(mfrow=c(2,2))
plot(ResG ~ FittedG, xlab="Fitted values", ylab="Residuals", main="Residuals vs. fitted")
abline(h=0) ## Homocedasticity
plot(log(df_anl$off) ~ FittedG, xlab="Fitted", ylab="Observed", main = "xyplot")
abline(0, 1) 
hist(ResG, main="Histogram of residuals", xlab="Residuals") ## Normality
boxplot(ResG,ylab = "Residuals")
```

Get temporal trends for on and off for each site and species.
```{r eval=do_eval}
get_coef_year <- function(df){df %>% filter(term == "year") %>% pull(estimate)}
get_p_year <- function(df){df %>% filter(term == "year") %>% pull(p.value)}

df_temporal <- df_temporal %>% 
  
  ## SOS
  mutate(linmod = purrr::map(data, ~lm(on ~ year, data = .))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_year_on = purrr::map_dbl(df_coef, ~get_coef_year(.))) %>% 
  mutate(p_value_year_on = purrr::map_dbl(df_coef, ~get_p_year(.))) %>% 
  select(-linmod, -df_coef) %>% 

  ## EOS
  mutate(linmod = purrr::map(data, ~lm(off ~ year, data = .))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_year_off = purrr::map_dbl(df_coef, ~get_coef_year(.))) %>% 
  mutate(p_value_year_off = purrr::map_dbl(df_coef, ~get_p_year(.))) %>% 
  select(-linmod, -df_coef)
  
save(df_temporal, file = "~/data/df_temporal.RData")
```

Temporal trends for on and off with LMMs
```{r}
length(unique(df_anl$id_species_site))
length(unique(df_anl$id_site))
length(unique(df_anl$species))
length(unique(df_anl$year))

Fit_Trend1 = lmer(off ~ year + (1|id_site/species) , data = df_anl, na.action = "na.exclude")
summary(Fit_Trend1)
r.squaredGLMM(Fit_Trend1)
plot(allEffects(Fit_IAV1))
Fit_Trend2 = lmer(on ~ year + (1|id_site/species) , data = df_anl, na.action = "na.exclude")
summary(Fit_Trend2)
r.squaredGLMM(Fit_Trend2)
plot(allEffects(Fit_Trend2))
```

Plot density of coefs
```{r}
load("data/df_temporal.RData")

df_temporal %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red", linetype = "dashed") +
  # geom_vline(xintercept = mean(df_temporal$coef_on, na.rm = TRUE), color = "red") +
  labs(title = "Slope of EOS ~ SOS", subtitle = paste(nrow(df_temporal), "sites and species"))

df_temporal %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_year_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_year_on, na.rm = TRUE), color = "red", linetype = "dashed") +
  # geom_vline(xintercept = mean(df_temporal$coef_year_on, na.rm = TRUE), color = "red") +
  labs(title = "Temporal trend of SOS", subtitle = paste(nrow(df_temporal), "sites and species"))

df_temporal %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_year_off, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_year_off, na.rm = TRUE), color = "red", linetype = "dashed") +
  # geom_vline(xintercept = mean(df_temporal$coef_year_off, na.rm = TRUE), color = "red") +
  labs(title = "Temporal trend of EOS", subtitle = paste(nrow(df_temporal), "sites and species"))

out <- df_temporal %>% 
  analyse_modobs2("coef_year_on", "coef_year_off")

out$gg +
  labs(title = "Trend in EOS vs. trend in SOS")
```

## Long-term

### By site

Get correlation between on and off for withing species-sites => long-term temporal variation (5 bins).
```{r eval=do_eval}
df_temporal_longterm <- df_anl %>% 
  # mutate(yearbin = ntile(year, n = 5)) %>% 
  mutate(yearbin = cut(year, breaks = 5, labels = FALSE)) %>%
  group_by(id_species_site, yearbin) %>%
  # group_by(id_site, yearbin) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(id_species_site) %>%
  # group_by(id_site) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_temporal_longterm, file = "~/data/df_temporal_longterm.RData")

## years in bins overview
df_anl %>% 
  # mutate(yearbin = ntile(year, n = 5)) %>% 
  mutate(yearbin = cut(year, breaks = 5, labels = FALSE)) %>%
  select(year, yearbin) %>% 
  ggplot(aes(year, yearbin)) +
  geom_point()
```

Long-term linear trends with LMMs
```{r}
length(unique(df_anl$id_species_site))
length(unique(df_anl$id_site))
length(unique(df_anl$species))
length(unique(df_anl$year))

Fit_LT = lmer(off ~ scale(on) + scale(year) + (1|id_site/species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT) # Removed interaction scale(on):scale(year) because was not significant
#plot(allEffects(Fit_LT))
plot_model(Fit_LT, type = "eff", terms = c("on","year[2020,1960,1890]"),title = "",axis.title = c("on","off"),legend.title = "year") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") 
r.squaredGLMM(Fit_LT)
```

Plot density of coefs
```{r}
load("data/df_temporal_longterm.RData")
df_temporal_longterm %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-20,20) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal_longterm$coef_on, na.rm = TRUE), color = "red", linetype = "dashed")
  # geom_vline(xintercept = mean(df_temporal_longterm$coef_on, na.rm = TRUE), color = "red")
```

### By site and species

Get correlation between on and off for withing species-sites => long-term temporal variation (5 bins).
```{r eval=do_eval}
df_temporal_longterm_sitepecies <- df_anl %>% 
  mutate(yearbin = cut(year, breaks = 5, labels = FALSE)) %>%
  group_by(id_species_site, yearbin) %>%
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(id_species_site) %>%
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_temporal_longterm_sitepecies, file = "data/df_temporal_longterm_sitepecies.RData")
```

Plot density of coefs
```{r}
load("data/df_temporal_longterm_sitepecies.RData")
df_temporal_longterm_sitepecies %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-20,20) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal_longterm$coef_on, na.rm = TRUE), color = "red", linetype = "dashed")
```

## Spatial

Get correlation between on and off for withing species-sites 
```{r eval=do_eval}
df_spatial <- df_anl %>% 
  group_by(id_site) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE))

save(df_spatial, file = "data/df_spatial.RData")
```

Plot correlation
```{r}
load("data/df_spatial.RData")
out_modobs <- df_spatial %>% analyse_modobs2("on", "off", type = "heat")
out_modobs$gg
```

Spatial by site with LMMs
```{r}
Fit_bySite = lmer(off ~ on + (1|id_site) , data = df_anl, na.action = "na.exclude")
summary(Fit_bySite)
plot(allEffects(Fit_bySite))
r.squaredGLMM(Fit_bySite)
```

## Spatial by species

Get correlation between on and off for withing species-sites 
```{r eval=do_eval}
df_spatial_species <- df_anl %>% 
  left_join(select(df_sub, id_species_site, species), by = c("id_species_site")) %>% 
  group_by(id_site, species) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(species) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

save(df_spatial_species, file = "data/df_spatial_species.RData")
df_spatial_species[[2]]
```

Spatial by species with LMMs
```{r}

df_spatial_sps <- df_anl %>% 
  group_by(id_site, species) %>% 
  summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE))

Fit_bySps = lmer(off ~ on + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps)
plot(allEffects(Fit_bySps))
r.squaredGLMM(Fit_bySps)
```

Plot density of coefs
```{r}
load("data/df_spatial_species.RData")
df_spatial_species %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram(binwidth = 0.5) +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_spatial_species$coef_on, na.rm = TRUE), color = "springgreen3")
```

## Long-term vs. IAV separation

DOESN'T WORK

```{r}
df_tmp <- df_anl %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(linmod_on = map(data, ~lm(on ~ year, data = .)),
         linmod_off = map(data, ~lm(off ~ year, data = .))) %>% 
  mutate(iav_on = map(linmod_on, "residuals"),
         iav_off = map(linmod_off, "residuals")) %>% 
  mutate(coef_on = map(linmod_on, "coefficients"),
         coef_off = map(linmod_off, "coefficients")) %>% 
  mutate(trend_on = map_dbl(coef_on, "year"),
         trend_off = map_dbl(coef_off, "year"))

## distribution of long-term trends for SOS and EOS
df_tmp %>% 
  pivot_longer(c(trend_on, trend_off), names_to = "pheno", values_to = "day") %>% 
  ggplot(aes(x = day, color = pheno)) +
  geom_density() +
  xlim(-2.5, 2.5)

## Long-term trend: no relationship
df_tmp %>% 
  filter(trend_on < 10 & trend_off < 10) %>% 
  ggplot(aes(trend_on, trend_off)) +
  geom_point()

out <- df_tmp %>% analyse_modobs2("trend_on", "trend_off", type = "hex")

df_iav <- df_tmp %>% 
  select(id_species_site, iav_on) %>% 
  unnest(iav_on) %>% 
  left_join(
    df_tmp %>% 
      select(id_species_site, iav_off) %>% 
      unnest(iav_off),
    by = "id_species_site"
  )

df_iav_nested <- df_iav %>%
  rename(on = iav_on, off = iav_off) %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% 
  mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% 
  select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  select(-summ)

df_iav_nested %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_iav_nested$coef_on, na.rm = TRUE), color = "red", linetype = "dashed") +
  labs(title = "Slope of IAV EOS ~ SOS", subtitle = paste(nrow(df_iav_nested), "sites and species"))
```

## Varying number of bins

Dependence of mean slope vs. aggregation time scale (number of bins). With Increasing aggregation (from right to left) reversal from a positive to a negative slope. Median slopes are plotted.

### Aggregating species

```{r eval=do_eval}
bins <- c(25, 19, 15, 11, 9, 7, 5, 3)

list_temporal_longterm <- list()
list_median_coef <- list()
for (ibreaks in bins){
  
  print(paste0("breaks_", ibreaks))

  list_temporal_longterm[[paste0("breaks_", ibreaks)]] <- df_anl %>%
    # mutate(yearbin = ntile(year, n = ibreaks)) %>%
    mutate(yearbin = cut(year, breaks = ibreaks, labels = FALSE)) %>%
    # group_by(id_species_site, yearbin) %>%   # DOESN'T WORK AS EXPECTED
    group_by(id_site, yearbin) %>%
    summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>%
    ungroup() %>%
    # group_by(id_species_site) %>%  # DOESN'T WORK AS EXPECTED
    group_by(id_site) %>%
    nest() %>%
    mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>%
    mutate(summ = purrr::map(linmod, ~summary(.))) %>%
    mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>%
    mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>%
    mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>%
    select(-linmod) %>%
    mutate(rsq = purrr::map_dbl(summ, "r.squared"),
           adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>%
    select(-summ)
  
  list_median_coef[[paste0("breaks_", ibreaks)]] <- median(list_temporal_longterm[[paste0("breaks_", ibreaks)]]$coef_on, na.rm = TRUE)
  
}
save(list_median_coef, file = "data/list_median_coef.RData")
save(list_temporal_longterm, file = "data/list_temporal_longterm.RData")

list_gg <- list()
for (ibreaks in bins){
  list_gg[[paste0("breaks_", ibreaks)]] <- list_temporal_longterm[[paste0("breaks_", ibreaks)]] %>% 
    # filter(p_value_on < 0.05) %>% 
    ggplot(aes(coef_on, ..density..)) +
    geom_histogram() +
    xlim(-10,10) +
    geom_vline(xintercept = 0, linetype = "dotted") +
    geom_vline(xintercept = median(list_temporal_longterm[[paste0("breaks_", ibreaks)]]$coef_on, na.rm = TRUE), color = "red", linetype = "dashed") +
    # geom_vline(xintercept = mean(list_temporal_longterm[[paste0("breaks_", ibreaks)]]$coef_on, na.rm = TRUE), color = "red")
    labs(title = paste("N bins", ibreaks))
}
save(list_gg, file = "data/list_gg.RData")
```

```{r}
bins <- c(25, 19, 15, 11, 9, 7, 5, 3)

load("data/list_gg.RData")
load("data/list_median_coef.RData")
load("data/list_temporal_longterm.RData")

# list_gg
vec_slopes <- c(list_median_coef %>% unlist(), out_modobs$results %>% pull(slope))
tibble(nbins =  c(bins, 1), median_slope = vec_slopes) %>% 
  ggplot(aes(nbins, median_slope)) +
  geom_point(size = 2) +
  geom_hline(yintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red") +
  geom_hline(yintercept = 0, linetype = "dotted")
```

## Summary

```{r}
df_temporal %>% 
  # filter(p_value_on < 0.05) %>% 
  ggplot(aes(coef_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red") +
  geom_vline(xintercept = median(df_spatial_species$coef_on, na.rm = TRUE), color = "springgreen3") +
  geom_vline(xintercept = out_modobs$df_metrics %>% filter(.metric == "slope") %>% pull(.estimate), color = "royalblue")
```
