---
title: "PEP725 explorations"
author: "Beni Stocker"
date: "12/2/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(rbeni)
library(lme4) #lmer
library(MuMIn) #r.squaredGLMM
library(lmerTest) #p-values
library(effects) #plot effects
library(sjPlot) #plot effects
do_eval <- FALSE
```

## Data loading and cleaning

Read the data complemented with drivers, obtained from Constantin Zohner (7.12.2020).
```{r}
df <- data.table::fread("~/data/pep/processed/DataMeta_3_Drivers_20_11_10.csv") %>% 
  as_tibble() %>% 
  rename(lon = LON, lat = LAT, year = YEAR, off = DoY_off, on = DoY_out, 
         anom_off_zani = autumn_anomaly, anom_on_zani = spring_anomaly, 
         species = Species, id_site = PEP_ID, id_species_site = timeseries) %>% 
  
  ## use the on-water-stressed version of A
  mutate(cA_tot = `cA_tot-w`)
```

Compare this to "my" dataset created in `pep.Rmd` (`df_anl`).
```{r}
df_mine <- read_csv("data/df_anl_zani.csv")

df_tmp <- df %>%
  left_join(df_mine, by = c("id_species_site", "year"))

df_tmp %>% analyse_modobs2("on.x", "on.y")
df_tmp %>% analyse_modobs2("off.x", "off.y")
```

Unfortunately, the SOS and EOS dates are not entirely identical for the two datasets. This could be due to applying the `distinct()` function somewhere along the process. There should really only be one single SOS and EOS data, resp., for each year, species, and site. Look into this again and find out what is done (and should be done) with duplicates.

```{r}
df_problem <- df_tmp %>%
  dplyr::filter(off.x != off.y) %>%
  dplyr::select(id_species_site, year, off_theirs = off.x, off_mine = off.y)
df_problem
```

Load the cleaned data frame before transforming into wider format.
```{r}
df_long <- data.table::fread("~/data/pep/processed/DataMeta_2_PhenologyObs_PEP725_CleanData.csv") %>%  #  "data/df_sub.csv"
  as_tibble() %>%
  mutate(pheno = ifelse(phenology == "leaf.out", "on", ifelse(phenology == "leaf.off", "off", NA))) %>%
  mutate(id_species_site = timeseries) %>%
  rename(day_correct = DoY, id_site = s_id) %>%
  dplyr::select(id_species_site, year, pheno, day_correct)

df_problem <- df_problem %>%
  left_join(df_long %>% dplyr::filter(pheno == "off"), by = c("id_species_site", "year")) %>%
  rowwise() %>%
  mutate(correct_mine = ifelse(off_mine == day_correct, TRUE, FALSE))

## check if any of my dates are not correct
any(!df_problem$correct_mine)
```

Mine is correct!

Select which one to use for further analysis.

<!-- Taking uncorrected Zani -->
<!-- ```{r} -->
<!-- df_anl <- df -->
<!-- ``` -->

Correct Zani data.
```{r}
df_correct <- data.table::fread("~/data/pep/processed/DataMeta_2_PhenologyObs_PEP725_CleanData.csv") %>%  #  "data/df_sub.csv"
  as_tibble() %>%
  mutate(pheno = ifelse(phenology == "leaf.out", "on", ifelse(phenology == "leaf.off", "off", NA))) %>%
  mutate(id_species_site = timeseries) %>%
  rename(day = DoY, id_site = s_id) %>%
  dplyr::select(id_species_site, year, pheno, day) %>% 
  pivot_wider(names_from = pheno, values_from = day)

df <- df %>%
  dplyr::rename(on_zani = on, off_zani = off) %>% 
  left_join(df_correct, by = c("id_species_site", "year"))

## some are actually missing - weird
sum(is.na(df$on))
sum(is.na(df$off))

df %>% 
  dplyr::filter(is.na(off))

## take zani data for these
df_anl <- df %>%
  rowwise() %>%
  mutate(on = ifelse(is.na(on), on_zani, on),
         off = ifelse(is.na(off), off_zani, off))

# ## or just drop them
# df_anl <- df %>% 
#   drop_na(on, off)
```

Add GPP from P-model siulations. XXX these are based on faulty EOS dates from Zani!!!
```{r}
load("data/df_pmodel.RData")

df_anl <- df_anl %>% 
  left_join(df_pmodel %>% 
              dplyr::select(id_species_site, out_pmodel) %>% 
              unnest(out_pmodel),
            by = c("id_species_site", "year"))

## add LUE
df_anl <- df_anl %>% 
  mutate(lue = gpp / apar)
```

Check consistency of P-model and LPJ-GUESS GPP.
```{r}
out <- df_anl %>% 
  analyse_modobs2("cA_tot", "gpp", type = "heat")
out$gg
```

Generally consistent except for a bunch of points where LPJ-GUESS GPP is very low. What's up with that?
```{r}
linmod <- lm(gpp ~ cA_tot, data = df_anl)

df_tmp <- df_anl %>% 
  mutate(diff = cA_tot - gpp) %>% 
  mutate(pdiff = diff / gpp)

df_sub <- df_tmp %>% 
  dplyr::filter(cA_tot < 800 & gpp > 100)

plot_map_simpl() +
  geom_point(
    data = df_sub, 
    aes(x = lon, y = lat, color = pdiff), 
    color = "red") +
  xlim(-10,30) + ylim(30, 60)

table(df_anl$species)/nrow(df_anl)
table(df_sub$species)/nrow(df_sub)
```

A lot of Larix decidua!


## Interannual

Get correlation between on and off for withing species-sites => interannual temporal variation.
```{r eval=do_eval}
get_coef_cA_tot <- function(df){df %>% dplyr::filter(term == "cA_tot") %>% pull(estimate)}
get_p_cA_tot <- function(df){df %>% dplyr::filter(term == "cA_tot") %>% pull(p.value)}

get_coef_gpp <- function(df){df %>% dplyr::filter(term == "gpp") %>% pull(estimate)}
get_p_gpp <- function(df){df %>% dplyr::filter(term == "gpp") %>% pull(p.value)}

get_coef_year <- function(df){df %>% dplyr::filter(term == "year") %>% pull(estimate)}
get_p_year <- function(df){df %>% dplyr::filter(term == "year") %>% pull(p.value)}

df_temporal <- df_anl %>% 
  
  ## for each species and site
  ungroup() %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  
  ## EOS ~ Assim
  mutate(linmod = purrr::map(data, ~lm(off ~ cA_tot, data = .))) %>% 
  mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_cA_tot = purrr::map_dbl(df_coef, ~get_coef_cA_tot(.))) %>% 
  mutate(p_value_cA_tot = purrr::map_dbl(df_coef, ~get_p_cA_tot(.))) %>% 
  dplyr::select(-linmod) %>% 
  mutate(rsq = purrr::map_dbl(summ, "r.squared"),
         adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  dplyr::select(-summ) %>% 
  
  # ## EOS ~ GPP P-model
  # mutate(linmod = purrr::map(data, ~lm(off ~ gpp, data = .))) %>% 
  # mutate(summ = purrr::map(linmod, ~summary(.))) %>% 
  # mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  # mutate(coef_gpp = purrr::map_dbl(df_coef, ~get_coef_gpp(.))) %>% 
  # mutate(p_value_gpp = purrr::map_dbl(df_coef, ~get_p_gpp(.))) %>% 
  # dplyr::select(-linmod) %>% 
  # mutate(rsq = purrr::map_dbl(summ, "r.squared"),
  #        adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% 
  # dplyr::select(-summ) %>% 

  ## SOS temporal trend
  mutate(linmod = purrr::map(data, ~lm(on ~ year, data = .))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_year_on = purrr::map_dbl(df_coef, ~get_coef_year(.))) %>% 
  mutate(p_value_year_on = purrr::map_dbl(df_coef, ~get_p_year(.))) %>% 
  dplyr::select(-linmod, -df_coef) %>% 

  ## EOS temporal trend
  mutate(linmod = purrr::map(data, ~lm(off ~ year, data = .))) %>% 
  mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% 
  mutate(coef_year_off = purrr::map_dbl(df_coef, ~get_coef_year(.))) %>% 
  mutate(p_value_year_off = purrr::map_dbl(df_coef, ~get_p_year(.))) %>% 
  dplyr::select(-linmod, -df_coef)
  
save(df_temporal, file = "~/data/df_interannual_assim.RData")
```

Plot distribution of slopes
```{r}
load("~/data/df_interannual_assim.RData")

df_temporal %>% 
  ggplot(aes(coef_year_on, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_year_on, na.rm = TRUE), color = "red", linetype = "dashed") +
  labs(title = "Temporal trend of SOS", subtitle = paste(nrow(df_temporal), "sites and species"))

df_temporal %>% 
  ggplot(aes(coef_year_off, ..density..)) +
  geom_histogram() +
  xlim(-2.5, 2.5) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_year_off, na.rm = TRUE), color = "red", linetype = "dashed") +
  labs(title = "Temporal trend of EOS", subtitle = paste(nrow(df_temporal), "sites and species"))

df_temporal %>% 
  ggplot(aes(coef_cA_tot, ..density..)) +
  geom_histogram() +
  xlim(-0.25, 0.25) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = median(df_temporal$coef_cA_tot, na.rm = TRUE), color = "red", linetype = "dashed") +
  labs(title = "Slope of EOS ~ cAtot", subtitle = paste(nrow(df_temporal), "sites and species"))

# df_temporal %>% 
#   ggplot(aes(coef_gpp, ..density..)) +
#   geom_histogram() +
#   xlim(-0.25, 0.25) +
#   geom_vline(xintercept = 0, linetype = "dotted") +
#   geom_vline(xintercept = median(df_temporal$coef_gpp, na.rm = TRUE), color = "red", linetype = "dashed") +
#   labs(title = "Slope of EOS ~ GPP P-model", subtitle = paste(nrow(df_temporal), "sites and species"))
```

Interannual temporal variation with LMMs
```{r}
# Both models are equivalent
Fit_IAV_cA_tot = lmer(off ~ cA_tot + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_cA_tot)
r.squaredGLMM(Fit_IAV_cA_tot)
plot(allEffects(Fit_IAV_cA_tot))

# Analysis of residuals
ResG <- residuals(Fit_IAV_cA_tot)
FittedG <- fitted(Fit_IAV_cA_tot)
par(mfrow=c(2,2))
plot(ResG ~ FittedG, xlab="Fitted values", ylab="Residuals", main="Residuals vs. fitted")
abline(h=0) ## Homocedasticity
plot(df_anl$off ~ FittedG, xlab="Fitted", ylab="Observed", main = "xyplot")
abline(0, 1) 
hist(ResG, main="Histogram of residuals", xlab="Residuals") ## Normality
boxplot(ResG,ylab = "Residuals")
```

With growing season index.
```{r}
Fit_IAV_cGSI = lmer(off ~ cGSI + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_cGSI)
r.squaredGLMM(Fit_IAV_cGSI)
plot(allEffects(Fit_IAV_cGSI))
```

With GPP from P-model
```{r}
Fit_IAV_gpp = lmer(off ~ gpp + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_gpp)
r.squaredGLMM(Fit_IAV_gpp)
plot(allEffects(Fit_IAV_gpp))
```

With APAR
```{r}
Fit_IAV_apar = lmer(off ~ apar + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_apar)
r.squaredGLMM(Fit_IAV_apar)
plot(allEffects(Fit_IAV_apar))
```

With LUE
```{r}
Fit_IAV_lue = lmer(off ~ lue + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_lue)
r.squaredGLMM(Fit_IAV_lue)
plot(allEffects(Fit_IAV_lue))
```

Weird pattern. Why is this?
```{r}
df_anl <- df_anl %>% 
  ungroup() %>% 
  mutate(res = residuals(Fit_IAV_cA_tot), fit = fitted(Fit_IAV_cA_tot))

## location of weird points
plot_map_simpl() +
  geom_point(
    data = df_anl %>% 
      dplyr::filter(fit > 310 & res < -40), 
    aes(x = lon, y = lat), 
    color = "red") +
  xlim(-10,30) + ylim(30, 60)

## distribution of A
ggplot() +
  geom_histogram(
    data = df_anl %>% 
      dplyr::filter(fit > 310 & res < -40),
    aes(x = cA_tot, y = ..density..),
    fill = "red",
    alpha = 0.5) +
  geom_histogram(
    data = df_anl %>% 
      dplyr::filter(!(fit > 310 & res < -40)),
    aes(x = cA_tot, y = ..density..),
    alpha = 0.5) +
  labs(title = "Assimilation")

## distribution of EOS
ggplot() +
  geom_histogram(
    data = df_anl %>% 
      dplyr::filter(fit > 310 & res < -40),
    aes(x = off, y = ..density..),
    fill = "red",
    alpha = 0.5) +
  geom_histogram(
    data = df_anl %>% 
      dplyr::filter(!(fit > 310 & res < -40)),
    aes(x = off, y = ..density..),
    alpha = 0.5) +
  labs(title = "EOS")

## distribution of years
ggplot() +
  geom_histogram(
    data = df_anl %>% 
      dplyr::filter(fit > 310 & res < -40),
    aes(x = year, y = ..density..),
    fill = "red",
    alpha = 0.5) +
  geom_histogram(
    data = df_anl %>% 
      dplyr::filter(!(fit > 310 & res < -40)),
    aes(x = year, y = ..density..),
    alpha = 0.5) +
  labs(title = "Year of observation")
```
These are only very few observations with weirdly low A.

They are clearly off the linear regression between A and EOS.
```{r}
df_anl <- df_anl %>% 
  mutate(off_fit = FittedG)

df_anl %>% 
  analyse_modobs2("off_fit", "off")
```

Identify them.
```{r}
df_anl <- df_anl %>% 
  # mutate(cluster = kmeans(select(., off, off_fit), centers = 2)$cluster) %>% 
  rowwise() %>% 
  mutate(cluster = ifelse( (abs(off - off_fit) > 43) & off_fit > 300, "bad", "good")) %>% 
  ungroup()

df_anl %>% 
  ggplot(aes(x = off_fit, y = off, color = cluster)) +
  geom_point()
```

Remove these sites. (NOT DONE HERE)
```{r eval=FALSE}
print(nrow(df_anl))
df_anl <- df_anl %>% 
  dplyr::filter(cluster == "good")
print(nrow(df_anl))
```

Run LMM for interannual temporal variation again.
```{r}
# Both models are equivalent
Fit_IAV_cA_tot = lmer(off ~ cA_tot + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_cA_tot)
r.squaredGLMM(Fit_IAV_cA_tot)
plot(allEffects(Fit_IAV_cA_tot))

Fit_IAV_cGSI = lmer(off ~ cGSI + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
summary(Fit_IAV_cGSI)
r.squaredGLMM(Fit_IAV_cGSI)
plot(allEffects(Fit_IAV_cGSI))
```

The residuals show a systematic relationship with years. As mixed effects models show, there is a long-term trend (relationship with 'year' as a fixed effect, see below).
```{r}
df_anl %>% 
  ungroup() %>% 
  mutate(res = residuals(Fit_IAV_cA_tot)) %>% 
  ggplot(aes(x = year, y = res)) +
  geom_point() +
  geom_smooth()
```

Still negative relationship!

We can re-create Fig. 1A in Zani et al. This shows correlation of points, pooling data from all species, and after "removing" the random factor 'site'. This essentially removes the spatial variation! This uses code adopted from `ModelAnalyses_1_Correlation&PathAnalyses_Figures1.R`.
```{r}
remove_randomfactor <- function(df){

  # fit model
  f1 <- lmer(off ~ cA_tot + (1|id_site), data = df)

  # Predict the data, removing the grouping value
  fix.pred1 <- predict(f1, re.form = NA)

  # To remove the random effects term and view the original spread of the data
  # with just the random noise added, add the residual error back
  # onto the predicted values
  y.adj1 <- fix.pred1 + resid(f1)

  # Store results
  df$off_norandom <- y.adj1

  return(df)
}

df_anl <- df_anl %>%
  group_by(species) %>%
  nest() %>%
  mutate(data = map(data, ~remove_randomfactor(.))) %>%
  unnest(data)

out <- df_anl %>%
  analyse_modobs2("cA_tot", "off_norandom", type = "heat")
out$gg +
  coord_cartesian(ylim = c(200,350))
```

Ok. This shows the same. The point is that any spatial variation is treated as a random effect and effectively removed by this method!

Do the same with GPP from P-model.
```{r}
remove_randomfactor <- function(df){

  df <- df %>% 
    dplyr::filter(!is.na(gpp))
  
  if (nrow(df) > 30){
    # fit model
    f1 <- lmer(off ~ gpp + (1|id_site), data = df)

    # Predict the data, removing the grouping value
    fix.pred1 <- predict(f1, re.form = NA)

    # To remove the random effects term and view the original spread of the data
    # with just the random noise added, add the residual error back
    # onto the predicted values
    y.adj1 <- fix.pred1 + resid(f1)

    # Store results
    df$off_norandom <- y.adj1 
  } else {
    df$off_norandom <- NA
  }

  return(df)
}

df_tmp <- df_anl %>%
  group_by(species) %>%
  nest() %>%
  mutate(data = map(data, ~remove_randomfactor(.))) %>%
  unnest(data)

out <- df_tmp %>%
  analyse_modobs2("gpp", "off_norandom", type = "heat")
out$gg +
  coord_cartesian(ylim = c(200,350))
```


<!-- Now, how does this relationship look like after aggregating to year bins? -->
<!-- ```{r} -->
<!-- remove_randomfactor_agg <- function(df, nbins){ -->

<!--   ## aggregate -->
<!--   df <- df %>%  -->
<!--     mutate(yearbin = cut(year, breaks = nbins, labels = FALSE)) %>% -->
<!--     group_by(id_site, yearbin) %>% -->
<!--     summarise(cA_tot = mean(cA_tot, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>%  -->
<!--     ungroup() -->

<!--   # fit model -->
<!--   f1 <- lmer(off ~ cA_tot + (1|id_site), data = df) -->

<!--   # Predict the data, removing the grouping value -->
<!--   fix.pred1 <- predict(f1, re.form = NA) -->

<!--   # To remove the random effects term and view the original spread of the data -->
<!--   # with just the random noise added, add the residual error back  -->
<!--   # onto the predicted values -->
<!--   y.adj1 <- fix.pred1 + resid(f1) -->

<!--   # Store results -->
<!--   df$off_norandom <- y.adj1 -->

<!--   return(df)  -->
<!-- } -->

<!-- df_anl_agg <- df_anl %>%  -->
<!--   group_by(species) %>%  -->
<!--   nest() %>%  -->
<!--   mutate(data = map(data, ~remove_randomfactor_agg(., nbins = 5))) %>%  -->
<!--   unnest(data) -->

<!-- out <- df_anl_agg %>%  -->
<!--   analyse_modobs2("cA_tot", "off_norandom", type = "heat") -->
<!-- out$gg + -->
<!--   coord_cartesian(ylim=c(200,350)) -->
<!-- ``` -->
<!-- Interesting. The negative relationship persists. -->

## Long-term

Separating the temporal trends from the relationship with A using LMMs. Without interactions between A and year.
```{r}
# SOS
Fit_LT_on = lmer(off ~ scale(on) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_on) 
plot(allEffects(Fit_LT_on))
plot_model(Fit_LT_on, type = "eff", terms = c("on","year[2020,1960,1890]"),title = "",axis.title = c("on","off"),legend.title = "year") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") 
r.squaredGLMM(Fit_LT_on)

# cA_tot: Interaction scale(cA_tot):scale(year) is significant and positive.
Fit_LT_cA_tot = lmer(off ~ scale(cA_tot) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_cA_tot) 
plot(allEffects(Fit_LT_cA_tot))
plot_model(Fit_LT_cA_tot, type = "eff", terms = c("cA_tot","year[2020,1960,1890]"),title = "",axis.title = c("cA_tot","off"),legend.title = "year") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") 
r.squaredGLMM(Fit_LT_cA_tot)

# cGSI
Fit_LT_cGSI = lmer(off ~ scale(cGSI) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_cGSI) 
plot(allEffects(Fit_LT_cGSI))
plot_model(Fit_LT_cGSI, type = "eff", terms = c("cGSI","year[2020,1960,1890]"),title = "",axis.title = c("cGSI","off"),legend.title = "year") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") 
r.squaredGLMM(Fit_LT_cGSI)

# P-model GPP
Fit_LT_gpp = lmer(off ~ scale(gpp) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_gpp) 
plot(allEffects(Fit_LT_gpp))
plot_model(Fit_LT_gpp, type = "eff", terms = c("gpp","year[2020,1960,1890]"),title = "",axis.title = c("gpp","off"),legend.title = "year") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") 
r.squaredGLMM(Fit_LT_gpp)
```

Separation of GPP in LUE and APAR.
```{r}
# P-model LUE and APAR (no time!)
Fit_LT_lue = lmer(off ~ scale(lue) + scale(apar) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_lue) 
plot(allEffects(Fit_LT_lue))
plot_model(Fit_LT_lue, type = "eff", terms = c("gpp", "lue[2020,1960,1890]"), title = "",axis.title = c("gpp","off"), legend.title = "lue") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "bottom") 
r.squaredGLMM(Fit_LT_lue)
```

Let's try to separate annual anomalies and site-level mean (of years 1970-2000).
```{r}
separate_anom <- function(df){
  df_mean <- df %>% 
    dplyr::filter(year %in% 1970:2000) %>% 
    summarise(mean_on = mean(on, na.rm = TRUE), 
              mean_off = mean(off, na.rm = TRUE), 
              mean_cA_tot = mean(cA_tot, na.rm = TRUE),
              mean_cGSI = mean(cGSI, na.rm = TRUE),
              mean_gpp = mean(gpp, na.rm = TRUE))
  df %>% 
    mutate(anom_on = on - df_mean$mean_on, 
           anom_off = off - df_mean$mean_off,
           mean_on = df_mean$mean_on, 
           mean_off = df_mean$mean_off,
           mean_cA_tot = df_mean$mean_cA_tot,
           anom_cA_tot = cA_tot - df_mean$mean_cA_tot,
           mean_cGSI = df_mean$mean_cGSI,
           anom_cGSI = cGSI - df_mean$mean_cGSI,
           mean_gpp = df_mean$mean_gpp,
           anom_gpp = gpp - df_mean$mean_gpp
           )
}
df_anl <- df_anl %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(data = map(data, ~separate_anom(.))) %>% 
  unnest(data)

## cA_tot
Fit_LT_anom = lmer(off ~ scale(mean_cA_tot) + scale(anom_cA_tot) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_anom)
plot(allEffects(Fit_LT_anom))
r.squaredGLMM(Fit_LT_anom)

## cGSI
Fit_LT_anom = lmer(off ~ scale(mean_cGSI) + scale(anom_cGSI) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_anom)
plot(allEffects(Fit_LT_anom))
r.squaredGLMM(Fit_LT_anom)

## gpp
Fit_LT_anom = lmer(off ~ scale(mean_gpp) + scale(anom_gpp) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")
summary(Fit_LT_anom)
plot(allEffects(Fit_LT_anom))
r.squaredGLMM(Fit_LT_anom)
```
Whatever I do, there is always this weird negative relationship between A and EOS at the spatial (between-site) scale.


<!-- ### By site -->

<!-- Get correlation between on and off for withing species-sites => long-term temporal variation (5 bins). -->
<!-- ```{r eval=do_eval} -->
<!-- df_temporal_longterm <- df_anl %>%  -->
<!--   mutate(yearbin = cut(year, breaks = 5, labels = FALSE)) %>% -->
<!--   group_by(id_site, yearbin) %>% -->
<!--   summarise(on = mean(on, na.rm = TRUE), cA_tot = mean(cA_tot, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   group_by(id_site) %>% -->
<!--   nest() %>%  -->
<!--   mutate(linmod = purrr::map(data, ~lm(off ~ cA_tot, data = .))) %>%  -->
<!--   mutate(summ = purrr::map(linmod, ~summary(.))) %>%  -->
<!--   mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>%  -->
<!--   mutate(coef_cA_tot = purrr::map_dbl(df_coef, ~get_coef_cA_tot(.))) %>%  -->
<!--   mutate(p_value_cA_tot = purrr::map_dbl(df_coef, ~get_p_cA_tot(.))) %>%  -->
<!--   dplyr::select(-linmod) %>%  -->
<!--   mutate(rsq = purrr::map_dbl(summ, "r.squared"), -->
<!--          adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>%  -->
<!--   dplyr::select(-summ) -->

<!-- save(df_temporal_longterm, file = "~/data/df_temporal_longterm_assim.RData") -->
<!-- ``` -->

<!-- Plot density of coefs -->
<!-- ```{r} -->
<!-- load("data/df_temporal_longterm_assim.RData") -->

<!-- df_temporal_longterm %>%  -->
<!--   ggplot(aes(coef_cA_tot, ..density..)) + -->
<!--   geom_histogram() + -->
<!--   xlim(-1,1) + -->
<!--   geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   geom_vline(xintercept = median(df_temporal_longterm$coef_cA_tot, na.rm = TRUE), color = "red", linetype = "dashed") -->
<!-- ``` -->



<!-- To separate the component of spatial variation, it's not as straight-forward as for long-term temporal variations (taking 'year' as a fixed effect in a linear mixed effects model). Let's just take the site-average SOS and the annual anomalies from it as two separate predictors, in addition to 'year' for the long-term trend. -->
<!-- ```{r} -->
<!-- separate_anom <- function(df){ -->
<!--   df_mean <- df %>%  -->
<!--     filter(year %in% 1970:2000) %>%  -->
<!--     summarise(mean_on = mean(on, na.rm = TRUE), mean_off = mean(off, na.rm = TRUE), mean_cA_tot = mean(cA_tot, na.rm = TRUE)) -->
<!--   df %>%  -->
<!--     mutate(anom_on = on - df_mean$mean_on,  -->
<!--            anom_off = off - df_mean$mean_off, -->
<!--            mean_on = df_mean$mean_on,  -->
<!--            mean_off = df_mean$mean_off, -->
<!--            mean_cA_tot = df_mean$mean_cA_tot, -->
<!--            anom_cA_tot = cA_tot - df_mean$mean_cA_tot, -->
<!--            ) -->
<!-- } -->
<!-- df_anl <- df_anl %>%  -->
<!--   group_by(id_species_site) %>%  -->
<!--   nest() %>%  -->
<!--   mutate(data = map(data, ~separate_anom(.))) %>%  -->
<!--   unnest(data) -->
<!-- ``` -->

<!-- Long-term linear trends, separated into site-specific-mean and anomalies. -->
<!-- ```{r} -->
<!-- Fit_LT_anom = lmer(off ~ scale(mean_cA_tot) + scale(anom_cA_tot) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude") -->
<!-- summary(Fit_LT_anom) -->
<!-- plot(allEffects(Fit_LT_anom)) -->
<!-- r.squaredGLMM(Fit_LT_anom) -->

<!-- ``` -->

<!-- ### By site and species -->

<!-- Get correlation between on and off for withing species-sites => long-term temporal variation (5 bins). -->
<!-- ```{r eval=do_eval} -->
<!-- df_temporal_longterm_sitepecies <- df_anl %>%  -->
<!--   mutate(yearbin = cut(year, breaks = 5, labels = FALSE)) %>% -->
<!--   group_by(id_species_site, yearbin) %>% -->
<!--   summarise(cA_tot = mean(cA_tot, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   group_by(id_species_site) %>% -->
<!--   nest() %>%  -->
<!--   mutate(linmod = purrr::map(data, ~lm(off ~ cA_tot, data = .))) %>%  -->
<!--   mutate(summ = purrr::map(linmod, ~summary(.))) %>%  -->
<!--   mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>%  -->
<!--   mutate(coef_cA_tot = purrr::map_dbl(df_coef, ~get_coef_cA_tot(.))) %>%  -->
<!--   mutate(p_value_cA_tot = purrr::map_dbl(df_coef, ~get_p_cA_tot(.))) %>%  -->
<!--   dplyr::select(-linmod) %>%  -->
<!--   mutate(rsq = purrr::map_dbl(summ, "r.squared"), -->
<!--          adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>%  -->
<!--   dplyr::select(-summ) -->

<!-- save(df_temporal_longterm_sitepecies, file = "data/df_temporal_longterm_sitepecies.RData") -->
<!-- ``` -->

<!-- Plot density of coefs -->
<!-- ```{r} -->
<!-- load("data/df_temporal_longterm_sitepecies.RData") -->
<!-- df_temporal_longterm_sitepecies %>%  -->
<!--   ggplot(aes(coef_cA_tot, ..density..)) + -->
<!--   geom_histogram() + -->
<!--   xlim(-1,1) + -->
<!--   geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   geom_vline(xintercept = median(df_temporal_longterm$coef_cA_tot, na.rm = TRUE), color = "red", linetype = "dashed") -->
<!-- ``` -->



## Spatial by species

Aggregate years, to get a "spatial" data frame, distinguishing species.
```{r}
df_spatial_sps <- df_anl %>% 
  group_by(id_site, species) %>% 
  summarise(cA_tot = mean(cA_tot, na.rm = TRUE), 
            off = mean(off, na.rm = TRUE), 
            cGSI = mean(cGSI, na.rm = TRUE),
            gpp = mean(gpp, na.rm = TRUE),
            on = mean(on, na.rm = TRUE))

Fit_bySps_cA_tot = lmer(off ~ cA_tot + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_cA_tot)
plot(allEffects(Fit_bySps_cA_tot))
r.squaredGLMM(Fit_bySps_cA_tot)

Fit_bySps_cGSI = lmer(off ~ cGSI + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_cGSI)
plot(allEffects(Fit_bySps_cGSI))
r.squaredGLMM(Fit_bySps_cGSI)

Fit_bySps_on = lmer(off ~ on + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_on)
plot(allEffects(Fit_bySps_on))
r.squaredGLMM(Fit_bySps_on)

Fit_bySps_gpp = lmer(off ~ gpp + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_gpp)
plot(allEffects(Fit_bySps_gpp))
r.squaredGLMM(Fit_bySps_gpp)
```
There is still a negative relationship between EOS and A, but not with GPP from P-model.

How does the relationship look by species? Not treated as a random effect, but done with separate linear regression models.
```{r}
list_linmod_cA_tot <- list()
list_summ_cA_tot <- list()
list_gg_cA_tot <- list()

list_linmod_cGSI <- list()
list_summ_cGSI <- list()
list_gg_cGSI <- list()

list_linmod_gpp <- list()
list_summ_gpp <- list()
list_gg_gpp <- list()

for (isp in unique(df_anl$species)){
  
  df_sub <- df_spatial_sps %>% 
    dplyr::filter(species == isp)

  list_linmod_cGSI[[isp]] <- lm(off ~ cGSI, data = df_sub, na.action = "na.exclude")
  list_summ_cGSI[[isp]] <- summary(list_linmod_cGSI[[isp]])
  list_gg_cGSI[[isp]] <- df_sub %>% analyse_modobs2("cGSI", "off", type = "heat")
  
  list_linmod_cA_tot[[isp]] <- lm(off ~ cA_tot, data = df_sub, na.action = "na.exclude")
  list_summ_cA_tot[[isp]] <- summary(list_linmod_cA_tot[[isp]])
  list_gg_cA_tot[[isp]] <- df_sub %>% analyse_modobs2("cA_tot", "off", type = "heat")
  
  list_linmod_gpp[[isp]] <- lm(off ~ gpp, data = df_sub, na.action = "na.exclude")
  list_summ_gpp[[isp]] <- summary(list_linmod_gpp[[isp]])
  list_gg_gpp[[isp]] <- df_sub %>% analyse_modobs2("gpp", "off", type = "heat")

}

purrr::map(list_gg_cA_tot, "gg")
purrr::map(list_gg_cGSI, "gg")
purrr::map(list_gg_gpp, "gg")
```



How are these drivers distributed in space?

Get correlation between on and off for withing species-sites 
```{r eval=do_eval}
getmostfrequent <- function(vec){names(sort(table(vec))[1])}
df_spatial <- df_anl %>% 
  group_by(id_site) %>% 
  summarise(cA_tot = mean(cA_tot, na.rm = TRUE), 
            off = mean(off, na.rm = TRUE), 
            cGSI = mean(cGSI, na.rm = TRUE), 
            on = mean(on, na.rm = TRUE),
            species = getmostfrequent(species)) %>% 
  left_join(select(df_anl, id_site, lon, lat) %>% unique(),
            by = "id_site")

save(df_spatial, file = "data/df_spatial_assim.RData")
```

Photosynthesis across space.
```{r}
plot_map_simpl(lonmin = 0, lonmax = 30, latmin = 40, latmax = 60) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = off)) +
  scale_color_viridis_c() +
  labs(title = "EOS")

plot_map_simpl(lonmin = 0, lonmax = 30, latmin = 40, latmax = 60) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = cGSI)) +
  scale_color_viridis_c() +
  labs(title = "GSI")

plot_map_simpl(lonmin = 0, lonmax = 30, latmin = 40, latmax = 60) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = cA_tot)) +
  scale_color_viridis_c() +
  labs(title = "Assimilation")
```

There are a few points with very low values for A (and probably late senescence) located in Switzerland - looks like in the Canton of Graubünden... (sorry for the ugly map)
```{r}
plot_map_simpl(lonmin = 5, lonmax = 12, latmin = 42, latmax = 50) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = off)) +
  scale_color_viridis_c() +
  labs(title = "EOS")

plot_map_simpl(lonmin = 5, lonmax = 12, latmin = 42, latmax = 50) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = cGSI)) +
  scale_color_viridis_c() +
  labs(title = "cGSI")

plot_map_simpl(lonmin = 5, lonmax = 12, latmin = 42, latmax = 50) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = cA_tot)) +
  scale_color_viridis_c() +
  labs(title = "Assimilation")

plot_map_simpl(lonmin = 5, lonmax = 12, latmin = 42, latmax = 50) +
  geom_point(data = df_spatial, aes(x = lon, y = lat, color = species)) +
  labs(title = "Species (most frequent at site)")
```

Yes. They are definitely in the mountains. And mostly Larix. Where else are Larix decidua distributed?
```{r}
plot_map_simpl(lonmin = 0, lonmax = 30, latmin = 40, latmax = 60) +
  geom_point(data = dplyr::filter(df_spatial, species == "Larix decidua"), aes(x = lon, y = lat)) +
  labs(title = "Larix decidua")
```

Ok. A bit everywhere.

Is there an interaction effect with elevation? See next section.


What's up with these sites that have a late season end and low A?
```{r}
df_spatial_sps <- df_spatial_sps %>% 
  ungroup() %>% 
  mutate(fit = fitted(Fit_bySps_cA_tot))

df_spatial_sps %>% 
  ggplot() +
  geom_line(aes(cA_tot, fit, color = species)) +
  geom_point(aes(cA_tot, off, color = species), alpha = 0.1)
```

Interaction effect with elevation?
```{r}
df_elv <- data.table::fread("~/data/pep/processed/DataMeta_2_PhenologyObs_PEP725_CleanData.csv") %>% 
  as_tibble() %>% 
  dplyr::select(id_site = s_id, alt, alt_dem) %>% 
  distinct() %>% 
  mutate(elv = ifelse(is.na(alt), alt_dem, alt)) %>% 
  dplyr::select(-alt, -alt_dem)

df_spatial_sps <- df_spatial_sps %>% 
  left_join(df_elv, by = c("id_site"))

Fit_bySps_cA_tot = lmer(off ~ cA_tot + elv + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_cA_tot)
plot(allEffects(Fit_bySps_cA_tot))
r.squaredGLMM(Fit_bySps_cA_tot)

Fit_bySps_cGSI = lmer(off ~ cGSI + elv + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_cGSI)
plot(allEffects(Fit_bySps_cGSI))
r.squaredGLMM(Fit_bySps_cGSI)

Fit_bySps_on = lmer(off ~ on + elv + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_on)
plot(allEffects(Fit_bySps_on))
r.squaredGLMM(Fit_bySps_on)
```









<!-- Get correlation between on and off for withing species-sites  -->
<!-- ```{r eval=do_eval} -->
<!-- df_spatial_species <- df_anl %>%  -->
<!--   group_by(id_site, species) %>%  -->
<!--   summarise(cA_tot = mean(cA_tot, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   group_by(species) %>%  -->
<!--   nest() %>%  -->
<!--   mutate(linmod = purrr::map(data, ~lm(off ~ cA_tot, data = .))) %>%  -->
<!--   mutate(summ = purrr::map(linmod, ~summary(.))) %>%  -->
<!--   mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>%  -->
<!--   mutate(coef_cA_tot = purrr::map_dbl(df_coef, ~get_coef_cA_tot(.))) %>%  -->
<!--   mutate(p_value_cA_tot = purrr::map_dbl(df_coef, ~get_p_cA_tot(.))) %>%  -->
<!--   dplyr::select(-linmod) %>%  -->
<!--   mutate(rsq = purrr::map_dbl(summ, "r.squared"), -->
<!--          adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>%  -->
<!--   dplyr::select(-summ) -->

<!-- save(df_spatial_species, file = "data/df_spatial_species_assim.RData") -->
<!-- ``` -->

<!-- Plot density of coefs. -->
<!-- ```{r} -->
<!-- load("data/df_spatial_species.RData") -->
<!-- df_spatial_species %>%  -->
<!--   ggplot(aes(coef_cA_tot, ..density..)) + -->
<!--   geom_histogram(binwidth = 0.5) + -->
<!--   xlim(-0.05, 0.05) + -->
<!--   geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   geom_vline(xintercept = median(df_spatial_species$coef_cA_tot, na.rm = TRUE), color = "springgreen3") -->
<!-- ``` -->

<!-- ## Varying number of bins -->

<!-- Dependence of mean slope vs. aggregation time scale (number of bins). With Increasing aggregation (from right to left) reversal from a positive to a negative slope. Median slopes are plotted. -->

<!-- ### Aggregating species -->

<!-- ```{r eval=do_eval} -->
<!-- bins <- c(25, 19, 15, 11, 9, 7, 5, 3) -->

<!-- list_temporal_longterm <- list() -->
<!-- list_median_coef <- list() -->
<!-- for (ibreaks in bins){ -->

<!--   print(paste0("breaks_", ibreaks)) -->

<!--   list_temporal_longterm[[paste0("breaks_", ibreaks)]] <- df_anl %>% -->
<!--     # mutate(yearbin = ntile(year, n = ibreaks)) %>% -->
<!--     mutate(yearbin = cut(year, breaks = ibreaks, labels = FALSE)) %>% -->
<!--     # group_by(id_species_site, yearbin) %>%   # DOESN'T WORK AS EXPECTED -->
<!--     group_by(id_site, yearbin) %>% -->
<!--     summarise(on = mean(on, na.rm = TRUE), off = mean(off, na.rm = TRUE)) %>% -->
<!--     ungroup() %>% -->
<!--     # group_by(id_species_site) %>%  # DOESN'T WORK AS EXPECTED -->
<!--     group_by(id_site) %>% -->
<!--     nest() %>% -->
<!--     mutate(linmod = purrr::map(data, ~lm(off ~ on, data = .))) %>% -->
<!--     mutate(summ = purrr::map(linmod, ~summary(.))) %>% -->
<!--     mutate(df_coef = purrr::map(linmod, ~tidy(.))) %>% -->
<!--     mutate(coef_on = purrr::map_dbl(df_coef, ~get_coef_on(.))) %>% -->
<!--     mutate(p_value_on = purrr::map_dbl(df_coef, ~get_p_on(.))) %>% -->
<!--     dplyr::select(-linmod) %>% -->
<!--     mutate(rsq = purrr::map_dbl(summ, "r.squared"), -->
<!--            adj_rsq = purrr::map_dbl(summ, "adj.r.squared")) %>% -->
<!--     dplyr::select(-summ) -->

<!--   list_median_coef[[paste0("breaks_", ibreaks)]] <- median(list_temporal_longterm[[paste0("breaks_", ibreaks)]]$coef_on, na.rm = TRUE) -->

<!-- } -->
<!-- save(list_median_coef, file = "data/list_median_coef.RData") -->
<!-- save(list_temporal_longterm, file = "data/list_temporal_longterm.RData") -->

<!-- list_gg <- list() -->
<!-- for (ibreaks in bins){ -->
<!--   list_gg[[paste0("breaks_", ibreaks)]] <- list_temporal_longterm[[paste0("breaks_", ibreaks)]] %>%  -->
<!--     # filter(p_value_on < 0.05) %>%  -->
<!--     ggplot(aes(coef_on, ..density..)) + -->
<!--     geom_histogram() + -->
<!--     xlim(-10,10) + -->
<!--     geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--     geom_vline(xintercept = median(list_temporal_longterm[[paste0("breaks_", ibreaks)]]$coef_on, na.rm = TRUE), color = "red", linetype = "dashed") + -->
<!--     # geom_vline(xintercept = mean(list_temporal_longterm[[paste0("breaks_", ibreaks)]]$coef_on, na.rm = TRUE), color = "red") -->
<!--     labs(title = paste("N bins", ibreaks)) -->
<!-- } -->
<!-- save(list_gg, file = "data/list_gg.RData") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- bins <- c(25, 19, 15, 11, 9, 7, 5, 3) -->

<!-- load("data/list_gg.RData") -->
<!-- load("data/list_median_coef.RData") -->
<!-- load("data/list_temporal_longterm.RData") -->

<!-- # list_gg -->
<!-- vec_slopes <- c(list_median_coef %>% unlist(), out_modobs$results %>% pull(slope)) -->
<!-- tibble(nbins =  c(bins, 1), median_slope = vec_slopes) %>%  -->
<!--   ggplot(aes(nbins, median_slope)) + -->
<!--   geom_point(size = 2) + -->
<!--   geom_hline(yintercept = median(df_temporal$coef_on, na.rm = TRUE), color = "red") + -->
<!--   geom_hline(yintercept = 0, linetype = "dotted") -->
<!-- ``` -->

<!-- ## Summary -->

<!-- ```{r} -->
<!-- df_temporal %>%  -->
<!--   ggplot(aes(coef_cA_tot, ..density..)) + -->
<!--   geom_histogram() + -->
<!--   xlim(-0.25, 0.25) + -->
<!--   geom_vline(xintercept = 0, linetype = "dotted") + -->
<!--   geom_vline(xintercept = median(df_temporal$coef_cA_tot, na.rm = TRUE), color = "red") + -->
<!--   geom_vline(xintercept = median(df_spatial_species$coef_cA_tot, na.rm = TRUE), color = "springgreen3") + -->
<!--   geom_vline(xintercept = out_modobs$df_metrics %>% dplyr::filter(.metric == "slope") %>% pull(.estimate), color = "royalblue") -->

<!-- ``` -->
