---
title: "P-model simulations for PEP725 data"
author: "Beni Stocker"
date: "12/2/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ingestr)
library(rbeni)
library(rsofun)
source("R/run_pmodel_pep.R")

do_eval <- TRUE
```

# Data preparation

Read the data complemented with drivers, obtained from Constantin Zohner (7.12.2020).
```{r}
df_zani <- data.table::fread("~/data/pep/processed/DataMeta_3_Drivers_20_11_10.csv") %>% 
  as_tibble() %>% 
  rename(lon = LON, lat = LAT, year = YEAR, off = DoY_off, on = DoY_out, 
         anom_off_zani = autumn_anomaly, anom_on_zani = spring_anomaly, 
         species = Species, s_id = PEP_ID, id_species_site = timeseries) %>% 
  
  ## use the on-water-stressed version of A
  mutate(cA_tot = `cA_tot-w`)
```

Read the dataset compiled by my implementation of the same criteria (as described by Zani et al., done in `pep.Rmd`)
```{r}
df_sub <- read_csv("data/df_sub.csv")

## there are some duplicates
dupl_on <- df_sub %>% 
  dplyr::select(s_id, id_species_site, species, year, pheno, day) %>% 
  dplyr::filter(pheno == "on") %>% 
  group_by(s_id, id_species_site, species, year) %>% 
  dplyr::filter(n() > 1)

df_problem <- dupl_on %>% 
  summarise(sdday = sd(day)) %>% 
  dplyr::filter(sdday > 0)

dupl_on %>% 
  dplyr::filter(id_species_site == df_problem$id_species_site[1] & year == df_problem$year[1])

## what does the original data say?
load("data/df_PEP725_alldata.RData") # loads 'df'
df_orig <- df %>% 
  dplyr::filter(phase_id %in% c(10, 11, 13, 95, 205)) %>% 
  mutate(pheno = ifelse(phase_id %in% c(10, 11, 13), "on", ifelse(phase_id %in% c(95, 205), "off", NA)))

## duplicates with inconsistent dates in the original data!
df_orig %>% 
  dplyr::filter(species == df_problem$species[1] & year == df_problem$year[1] & s_id == df_problem$s_id[1])

## take the mean across inconsistent duplicates. 3,223,360 -> 2,987,845 data points.
df_orig <- df_orig %>% 
  dplyr::select(s_id, species, year, pheno, day) %>% 
  group_by(s_id, species, year, pheno) %>% 
  summarise(day = round(mean(day), 0))

df_orig_wide <- df_orig %>% 
   pivot_wider(names_from = "pheno", values_from = "day")
```

Correct Zani pheno dates
```{r}
df_zani <- df_zani %>% 
  rename(on_zani = on, off_zani = off) %>% 
  left_join(df_orig_wide, by = c("s_id", "species", "year"))

out <- df_zani %>% analyse_modobs2("off", "off_zani", type = "heat")
out$gg
```

Take my data subset and overwrite pheno dates with the consolidated ones (mean across duplicates).
```{r}
df_mine <- df_sub %>% 
  dplyr::select(s_id, lon, lat, alt, alt_dem, species, year, pheno) %>% 
  distinct() %>% 
  left_join(df_orig, by = c("s_id", "species", "year", "pheno")) %>% 
  pivot_wider(names_from = "pheno", values_from = "day") %>% 
  drop_na(on, off)
```

## Overview of sites.

```{r}
df_sites_mine <- df_mine %>% 
  dplyr::select(s_id, lon, lat) %>% 
  distinct()

df_sites_zani <- df_zani %>% 
  dplyr::select(s_id, lon, lat) %>% 
  distinct()

missing_zani <- df_sites_zani %>% 
  dplyr::filter(!(s_id %in% df_sites_mine$s_id))

missing_mine <- df_sites_mine %>% 
  dplyr::filter(!(s_id %in% df_sites_zani$s_id))

plot_map_simpl(lonmin = -10, lonmax = 40, latmin = 30, latmax = 70) +
  geom_point(data = df_sites_mine, aes(x = lon, y = lat), col = "red")

plot_map_simpl(lonmin = -10, lonmax = 40, latmin = 30, latmax = 70) +
  geom_point(data = df_sites_zani, aes(x = lon, y = lat), col = "red")
```
## Bin data

Bin the data into 0.5 deg grid points.

Zani data.
```{r}
## bin
dlon <- 0.5
dlat <- 0.5
lon_breaks <- seq(from = floor(min(df_zani$lon)), to = ceiling(max(df_zani$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df_zani$lat)), to = ceiling(max(df_zani$lat)), by = dlat)

df_zani <- df_zani %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2)

df_zani <- df_zani %>% 
  group_by(lon_mid, lat_mid) %>% 
  drop_na(lon_mid, lat_mid)

save(df_zani, file = "data/df_zani_corrected.RData")
```

My data.
```{r}
## bin
dlon <- 0.5
dlat <- 0.5
lon_breaks <- seq(from = floor(min(df_mine$lon)), to = ceiling(max(df_mine$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df_mine$lat)), to = ceiling(max(df_mine$lat)), by = dlat)

df_mine <- df_mine %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2)

df_mine <- df_mine %>% 
  group_by(lon_mid, lat_mid) %>% 
  drop_na(lon_mid, lat_mid)

save(df_mine, file = "data/df_mine.RData")
```


Get all gridcells with at least one site.
```{r}
df_zani_cells <- df_zani %>% 
  ungroup() %>% 
  dplyr::select(lon = lon_mid, lat = lat_mid) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(year_start = 1979, year_end = 2018) %>% 
  mutate(lon_mid = lon, lat_mid = lat) %>% 
  mutate(sitename = paste0("icell_", 1:n()))

save(df_zani_cells, file = "data/df_zani_cells.RData")

df_mine_cells <- df_mine %>% 
  ungroup() %>% 
  dplyr::select(lon = lon_mid, lat = lat_mid) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(year_start = 1979, year_end = 2018) %>% 
  mutate(lon_mid = lon, lat_mid = lat) %>% 
  mutate(sitename = paste0("icell_", 1:n()))

save(df_mine_cells, file = "data/df_mine_cells.RData")

  
## complement with elevation data
df_zani <- data.table::fread("~/data/pep/processed/DataMeta_2_PhenologyObs_PEP725_CleanData.csv") %>% 
  as_tibble() %>% 
  dplyr::select(s_id, alt, alt_dem, lon, lat) %>% 
  distinct() %>% 
  mutate(elv = ifelse(is.na(alt), alt_dem, alt)) %>% 
  dplyr::select(-alt, -alt_dem) %>% 
  right_join(df_zani, by = c("lon", "lat", "s_id"))

df_mine <- df_mine %>% 
  mutate(elv = ifelse(is.na(alt), alt_dem, alt))

## fill missing elv by mean
mean_elv <- mean(df_zani$elv, na.rm = TRUE)
df_zani <- df_zani %>% 
  rowwise() %>% 
  mutate(elv = ifelse(is.na(elv), mean_elv, elv)) %>% 
  ungroup()
df_mine <- df_mine %>% 
  rowwise() %>% 
  mutate(elv = ifelse(is.na(elv), mean_elv, elv)) %>% 
  ungroup()
```

## Get data.

This is done now by `rscript_get_watch_cru.R` and `submit_get_watch_cru.sh`

Combine the two meteo data frames into one, containing `ccov` (cloud cover) from CRU and all other variables from FLUXNET.
```{r}
load("data/ddf_watch_zani.RData")
load("data/ddf_cru_zani.RData")

ddf_meteo_zani <- ddf_watch_zani %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru_zani %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest()
```

```{r}
load("data/ddf_watch_mine.RData")
load("data/ddf_cru_mine.RData")

ddf_meteo_mine <- ddf_watch_mine %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru_mine %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest()
```

CO2
```{r}
df_co2 <- read_csv("~/data/co2/cCO2_rcp85_const850-1765.csv")
```

## Run model

```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)

## calibrated parameters for v3.0 (see https://rpubs.com/stineb/rsofun_benchmark_v30)
params_modl <- list(
	kphio           = 0.09423773,
	soilm_par_a     = 0.33349283,
	soilm_par_b     = 1.45602286,
	vpdstress_par_a = 9999,
	vpdstress_par_b = 9999,
	vpdstress_par_m = 9999
	)

df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Zani

```{r}  
df_pmodel_zani <- df_zani %>% 
  
  ## nest by site-species (timeseries)
  group_by(id_species_site, lon_mid, lat_mid) %>% 
  nest() %>% 
  
  ## add id_cell column
  left_join(df_zani_cells %>% 
              dplyr::select(-lon, -lat),
            by = c("lon_mid", "lat_mid")) %>% 
  ungroup() %>% 
  
  ## run p-model
  mutate(out_pmodel = purrr::map2(data, sitename, 
                                  ~run_pmodel_pep(
                                    .x,
                                    ddf_meteo_zani %>% 
                                      dplyr::filter(sitename == .y) %>% 
                                      unnest(data),
                                    df_co2,
                                    df_zani_cells %>% 
                                      dplyr::filter(sitename == .y),
                                    params_siml, 
                                     params_modl, 
                                    df_soiltexture 
                                  ))) %>% 
  dplyr::select(-data)

save(df_pmodel_zani, file = "data/df_pmodel_zani.RData")
```

Compare to LPJ-GUESS from Zani et al.
```{r}
df_comb <- df_pmodel_zani %>% 
  unnest(out_pmodel) %>% 
  right_join(df_zani, by = c("id_species_site", "year")) %>% 
  mutate(gpp_net = gpp - rd)

df_comb %>% analyse_modobs2("cA_tot", "gpp_net", type = "heat")
```


<!-- Test -->
<!-- ```{r} -->
<!-- df_pmodel$out_pmodel[[2]] %>%  -->
<!--   dplyr::filter(year(date) %in% 1977:1980) %>%  -->
<!--   # slice(1:1000) %>%  -->
<!--   ggplot(aes(date, gpp)) +  -->
<!--   geom_line() -->

<!-- test <- df_pmodel$out_pmodel[[1]] %>%  -->
<!--   mutate(year = lubridate::year(date)) %>%  -->
<!--   group_by(year) %>%  -->
<!--   summarise_if(is.numeric, max, na.rm = TRUE) -->

<!-- plot(test$gpp, type = "l") -->
<!-- ``` -->

<!-- Weirdly low gpp values! -->
<!-- ```{r} -->
<!-- ## xxx debug -->
<!-- df_pheno = df_pmodel$data[[1]] -->
<!-- df_forcing = ddf_meteo %>% dplyr::filter(sitename == "icell_1") %>% unnest(data) -->
<!-- df_siteinfo = df_sites %>% dplyr::filter(sitename == "icell_1") -->

<!-- ## compare to fluxnet data from germany -->
<!-- load("~/data/rsofun_benchmarking/df_drivers_fluxnet2015.Rdata") -->
<!-- tmp <- df_drivers_fluxnet2015 %>%  -->
<!--   dplyr::filter(str_detect(sitename, "DE")) %>%  -->
<!--   dplyr::select(forcing) %>%  -->
<!--   unnest(forcing) %>%  -->
<!--   mutate(prec = prec / (60*60*24), ppfd = ppfd / (60*60*24)) -->

<!-- ggplot() + -->
<!--   geom_histogram(data = tmp, aes(x = prec, y = ..density..), alpha = 0.5) + -->
<!--   geom_histogram(data = df_forcing, aes(x = prec, y = ..density..), alpha = 0.5, fill = "red") +  -->
<!--   xlim(-0.0001, 0.0005) -->

<!-- ggplot() + -->
<!--   geom_histogram(data = tmp, aes(x = ppfd, y = ..density..), alpha = 0.5) + -->
<!--   geom_histogram(data = df_forcing, aes(x = ppfd, y = ..density..), alpha = 0.5, fill = "red") -->

<!-- skim(tmp) -->
<!-- skim(df_forcing) -->
<!-- ``` -->



