---
title: "Spatial sampling of PEP725 data"
author: "Beni Stocker"
date: "12/2/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(rbeni)
library(lme4) #lmer
library(MuMIn) #r.squaredGLMM
library(lmerTest) #p-values
library(effects) #plot effects
library(sjPlot) #plot effects
do_eval <- TRUE
```

### Analysis

Read the data complemented with drivers, obtained from Constantin Zohner (7.12.2020).
```{r}
df <- data.table::fread("~/data/pep/processed/DataMeta_3_Drivers_20_11_10.csv") %>% 
  as_tibble() %>% 
  rename(lon = LON, lat = LAT, year = YEAR, off = DoY_off, on = DoY_out, 
         anom_off_zani = autumn_anomaly, anom_on_zani = spring_anomaly, 
         species = Species, id_site = PEP_ID, id_species_site = timeseries) %>% 
  
  ## use the on-water-stressed version of A
  mutate(cA_tot = `cA_tot-w`)
```

Bin the data into 0.5 deg grid points.
```{r}
## bin
dlon <- 1
dlat <- 1
lon_breaks <- seq(from = floor(min(df$lon)), to = ceiling(max(df$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df$lat)), to = ceiling(max(df$lat)), by = dlat)

df <- df %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2)

#View(select(df, lon, lat, ilon, ilat, lon_lower, lon_upper, lat_lower, lat_upper, lon_mid, lat_mid))

df <- df %>% 
  group_by(lon_mid, lat_mid) %>% 
  drop_na(lon_mid, lat_mid)
```

The distribution of number of sites per gridcell.
```{r}
tmp <- df %>% 
  select(id_site, lon_mid, lat_mid) %>% 
  distinct() %>% 
  group_by(lon_mid, lat_mid) %>% 
  summarise(n = n())

tmp %>% 
  ggplot(aes(x = n, y = ..count..)) +
  geom_histogram() +
  labs(title = "Number of sites in half-degree gridcells")
```

Sample one site per gridcell.
```{r}
set.seed(189)
df_sites_sampled <- df %>% 
  select(id_site, lon_mid, lat_mid) %>% 
  distinct() %>% 
  group_by(lon_mid, lat_mid) %>% 
  sample_n(1)
```

Subset original data frame based on this selection.
```{r}
df_sampled <- df %>% 
  filter(id_site %in% df_sites_sampled$id_site)
```

Quick check: does it still have the negative relationship between A and EOS at the spatial scale?
```{r}
df_spatial_sps <- df_sampled %>% 
  group_by(id_site, species) %>% 
  summarise(cA_tot = mean(cA_tot, na.rm = TRUE), off = mean(off, na.rm = TRUE), cGSI = mean(cGSI, na.rm = TRUE), on = mean(on, na.rm = TRUE))

Fit_bySps_cA_tot = lmer(off ~ cA_tot + (1|species), data = df_spatial_sps, na.action = "na.exclude")
summary(Fit_bySps_cA_tot)
plot(allEffects(Fit_bySps_cA_tot))
r.squaredGLMM(Fit_bySps_cA_tot)

df_spatial_sps %>% 
  analyse_modobs2("cA_tot", "off")
```
Is still there, but only weak.

```{r}
plot_map_simpl() +
  geom_point(
    data = df_sampled, 
    aes(x = lon, y = lat), 
    color = "red") +
  xlim(-10,30) + ylim(30, 60)
```