---
title: "Zani et al. explorations"
author: "Beni Stocker"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(rbeni)
do_eval <- FALSE
```

## Data cleaning

**Code and text explanations adopted (and adjusted for slightly different file formats) from `DataMeta_1_PhenoData_Cleaning.R`.**

Following ref(43), we removed 
(i) individual time-series with less than 15 years of leaf-out and leaf senescence observations for the same year, 
(ii) dates deviating from an individual's median more than 3 times the median absolute deviation, 
(iii) time-series for which the standard deviation of phenological observations across years was higher than 15 for leaf flushing and 20 for leaf senescence. 
The thresholds differed because the mean absolute deviation for leaf-out time-series (8.8) was lower than that for leaf senescence (10.0). 

Import data. This is done slightly differently here, because we I obtained data separated into multiple files.
```{r eval=do_eval}
vec_files <- list.files("~/data/pep/PEP725_2020-12/", pattern = "pep725_2020-12_", full.names = TRUE)
list_df <- map(as.list(vec_files), ~read.delim(., sep = ";"))
names(list_df) <- vec_files

df <- list_df %>% 
  bind_rows(.id = "filnam") %>% 
  as_tibble() %>%  
  mutate(date = ymd(date))

save(df, file = "data/df_pep725.RData")
```

```{r}
load("data/df_pep725.RData")
```


### Number of observations per species

Eliminate species with less than 3000 observations.
```{r}
species_retain <- df %>% 
  group_by(species) %>% 
  summarise(n_obs = n()) %>% 
  filter(n_obs >= 3000) %>% 
  pull(species)

df_sub <- df %>% 
  filter(species %in% species_retain)
```

This reduces original dataset to (fraction):
```{r}
nrow(df_sub)/nrow(df)
```

### Relevant phenological dates

Keep only records on leaf flushing and leaf senescence dates
From PEP725_BBCH.csv (with definition of the BBCH Code) from the PEP725 Dataset
Spring phenology codes:

- 10	First leaves seperated (mouse ear)
- 11	Leaf unfolding (first visible leaf stalk)

Autumn phenology codes
- 94	Autumnal coloring of leaves (50%)
- 95	50 % autumnal leaf fall

**There seems to be some disaccordance between the code key above and definitions given on the respective page online [here](http://www.pep725.eu/pep725_phase.php) and the selection indicated in the SI:**

- **Code 94 is not in the dataset. Taking 205 instead ("autumnal colouration >=50%")**
- **Code 13 described as being used in SI, but not in code.**

```{r}
df_sub <- df_sub %>% 
  filter(phase_id %in% c(10, 11, 95, 205)) %>%    # include 13?
  mutate(phenology = ifelse(phase_id %in% c(10, 11), "leaf.out", ifelse(phase_id %in% c(95, 205), "leaf.off", NA)))
```

Create timeseries identifier (site x species).

**In the present dataset, site ID is given by** `s_id`.

```{r}
df_species_sites <- df_sub %>% 
  select(s_id, species) %>% 
  distinct() %>% 
  mutate(timeseries = paste0(s_id, "_", species))  # column corresponds to id_species_site

df_sub <- df_sub %>% 
  left_join(df_species_sites, by = c("s_id", "species"))
```

Keep only timeseries with observations for both spring and autumn leaf phenology.

**My tidyverse code doesn't work here because there are duplicates in terms of columns `c(timeseries, phenology, day)`!**

```{r eval=FALSE}
df_retain <- df_sub %>% 
  select(timeseries, phenology, day) %>% 
  pivot_wider(names_from = "phenology", values_from = "day")
```

Using Zani's code instead.
```{r}
pheno_out.df <- df_sub %>% 
  filter(phenology == "leaf.out")
pheno_off.df <- df_sub %>% 
  filter(phenology == "leaf.off")

ts_common <- intersect(pheno_out.df$timeseries, pheno_off.df$timeseries)
pheno_out.df <- pheno_out.df[pheno_out.df$timeseries %in% ts_common,]
pheno_off.df <- pheno_off.df[pheno_off.df$timeseries %in% ts_common,]    # XXX here, Zani's code makes pheno_out.df <- ... THIS IS A BUG! CORRECTED HERE. XXX
```

Save unique site locations in CSV (for other uses).
```{r}
df_siteinfo <- df_sub %>% 
  select(s_id, lon, lat, year) %>% 
  group_by(s_id) %>% 
  summarise(year_start = min(year), year_end = max(year), lon = mean(lon), lat = mean(lat), var_lon = var(lon), var_lat = var(lat)) %>% 
  distinct()

df_siteinfo %>% 
  write_csv("data/siteinfo_pep.csv")
```

Apply remaining data cleaning steps, using function with code adopted from Zani's code.
```{r}
pheno_clean.df <- clean_zani(pheno_out.df, pheno_off.df, ts_common)
write.table(pheno_clean.df, "data/DataMeta_2_PhenologyObs_PEP725_CleanData.csv", sep=";", row.names = FALSE)
```
