---
title: "Comparison of modelled photosynthesis in PEP725 simulations"
author: "Beni Stocker"
date: "12/2/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ingestr) # see github of stineb
library(rbeni)   # see github of stineb
library(rsofun)  # see github of stineb
source("R/run_pmodel_pep.R")

do_eval <- TRUE
```

This is to run models at a few sites and compare results for daily photosynthesis obtained by different models: LPJ implementation by Zani et al., P-model implementation by Constantin and by Beni.

# Data preparation

Read the data complemented with drivers, obtained from Constantin Zohner (7.12.2020).
```{r}
df_zanitest <- data.table::fread("~/data/pep/processed/DataMeta_3_Drivers_20_11_10.csv") %>% 
  as_tibble() %>% 
  rename(lon = LON, lat = LAT, year = YEAR, off = DoY_off, on = DoY_out, 
         anom_off_zani = autumn_anomaly, anom_on_zani = spring_anomaly, 
         species = Species, s_id = PEP_ID, id_species_site = timeseries) 
  
  # ## use the on-water-stressed version of A
  # mutate(cA_tot = `cA_tot-w`)
```


## Bin data

Bin the data by lon-lat grid points.
```{r}
## bin
dlon <- 2
dlat <- 2
lon_breaks <- seq(from = floor(min(df_zanitest$lon)), to = ceiling(max(df_zanitest$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df_zanitest$lat)), to = ceiling(max(df_zanitest$lat)), by = dlat)

df_zanitest <- df_zanitest %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2)

df_zanitest <- df_zanitest %>% 
  group_by(lon_mid, lat_mid) %>% 
  drop_na(lon_mid, lat_mid)

#save(df_zanitest, file = "data/df_zanitest_corrected.RData")
```


Get all gridcells with at least one site.
```{r}
df_zanitest_cells <- df_zanitest %>% 
  ungroup() %>% 
  dplyr::select(lon = lon_mid, lat = lat_mid) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(year_start = 1979, year_end = 2018) %>% 
  mutate(lon_mid = lon, lat_mid = lat) %>% 
  mutate(cellname = paste0("icell_", 1:n()))

save(df_zanitest_cells, file = "data/df_zanitest_cells.RData")

## complement with elevation data
df_zanitest <- data.table::fread("~/data/pep/processed/DataMeta_2_PhenologyObs_PEP725_CleanData.csv") %>% 
  as_tibble() %>% 
  dplyr::select(s_id, alt, alt_dem, lon, lat) %>% 
  distinct() %>% 
  mutate(elv = ifelse(is.na(alt), alt_dem, alt)) %>% 
  dplyr::select(-alt, -alt_dem) %>% 
  right_join(df_zanitest, by = c("lon", "lat", "s_id"))

## fill missing elv by mean
mean_elv <- mean(df_zanitest$elv, na.rm = TRUE)
df_zanitest <- df_zanitest %>% 
  rowwise() %>% 
  mutate(elv = ifelse(is.na(elv), mean_elv, elv)) %>% 
  ungroup()
```

Sample one site per gridcell.
```{r}
## consider only sites with sufficiently long time series
df_nyears <- df_zanitest %>% 
  group_by(id_species_site) %>% 
  summarise(nyears = n())

df_nyears %>% 
  ggplot(aes(x = nyears, y = ..count..)) +
  geom_histogram()

set.seed(1982)
df_tseries_sampled <- df_zanitest %>% 
  
  # ## take only those that cover >20 years
  # left_join(df_nyears, by = c("id_species_site")) %>% 
  # dplyr::filter(nyears > 25) %>% 
  
  ## sample one site per spatial gridcell (to get good spatial coverage)
  dplyr::select(id_species_site, lon, lat, elv, lon_mid, lat_mid) %>%
  distinct() %>%
  group_by(lon_mid, lat_mid) %>% 
  sample_n(1) %>% 
  
  ## add cell name info
  left_join(df_zanitest_cells %>% dplyr::select(lon_mid, lat_mid, cellname),
            by = c("lon_mid", "lat_mid"))

## subset data to sampled time series (33)
df_zanitest_sub <- df_zanitest %>% 
  dplyr::filter(id_species_site %in% df_tseries_sampled$id_species_site)

## determine start and end year of sampled t series
df_tseries_sampled <- df_tseries_sampled %>% 
  left_join(
    df_zanitest_sub %>% 
      group_by(id_species_site) %>% 
      summarise(year_start = min(year), year_end = max(year)),
    by = "id_species_site"
  )
```

Overview of sampled sites.
```{r}
plot_map_simpl() +
  geom_point(
    data = df_tseries_sampled, 
    aes(x = lon, y = lat), 
    color = "red") +
  xlim(-10,30) + ylim(30, 60)
```


## Get forcing data

Re-do this because the longitude and latitude info was not recorded.
```{r eval=FALSE}
ddf_watch_zanitest <- ingest(
  siteinfo = df_tseries_sampled %>% 
    rename(sitename = id_species_site) %>% 
    mutate(year_start_tseries = year_start,
           year_end_tseries = year_end) %>% 
    mutate(year_start = max(1979, year_start),
           year_end = max(1979, year_end)),
  source    = "watch_wfdei",
  getvars   = c("temp", "prec", "ppfd", "vpd", "patm"),
  dir       = "~/data/watch_wfdei/"  # adjust this with your local path
  ) %>% 
  rename(id_species_site = sitename)

save(ddf_watch_zanitest, file = "data/ddf_watch_zanitest.RData")

ddf_cru_zanitest <- ingest(
  siteinfo = df_tseries_sampled %>% 
    rename(sitename = id_species_site) %>% 
    mutate(year_start_tseries = year_start,
           year_end_tseries = year_end) %>% 
    mutate(year_start = max(1979, year_start),
           year_end = max(1979, year_end)),
  source    = "cru",
  getvars   = "ccov",
  dir       = "~/data/cru/ts_4.01/"  # adjust this with your local path
  ) %>% 
  rename(id_species_site = sitename)

save(ddf_cru_zanitest, file = "data/ddf_cru_zanitest.RData")
```

Combine the two meteo data frames into one, containing `ccov` (cloud cover) from CRU and all other variables from FLUXNET.
```{r}
load("data/ddf_watch_zanitest.RData")
load("data/ddf_cru_zanitest.RData")

ddf_meteo_zanitest <- ddf_watch_zanitest %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru_zanitest %>% 
      tidyr::unnest(data),
    by = c("id_species_site", "date")
  ) %>% 
  group_by(id_species_site) %>% 
  tidyr::nest() %>% 
  left_join(df_tseries_sampled, by = "id_species_site")
```

CO2
```{r}
df_co2 <- read_csv("~/data/co2/cCO2_rcp85_const850-1765.csv")
```

## Run model

```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)

## calibrated parameters for v3.0 (see https://rpubs.com/stineb/rsofun_benchmark_v30)
params_modl <- list(
	kphio           = 0.09423773,
	soilm_par_a     = 0.33349283,
	soilm_par_b     = 1.45602286,
	vpdstress_par_a = 9999,
	vpdstress_par_b = 9999,
	vpdstress_par_m = 9999
	)

df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```


```{r}  
df_pmodel_zanitest <- df_zanitest_sub %>% 
  
  ## nest by site-species (timeseries)
  group_by(id_species_site, lon_mid, lat_mid) %>% 
  nest() %>% 
  
  ## add id_cell column
  left_join(df_zanitest_cells %>%   # this must contain the same info for 'sitename' as is used for creating forcing dataset
              dplyr::select(-lon, -lat),
            by = c("lon_mid", "lat_mid")) %>% 
  ungroup() %>% 
  
  ## run p-model
  mutate(out_pmodel = purrr::map2(data, cellname, 
                                  ~run_pmodel_pep(
                                    .x,
                                    ddf_meteo_zanitest %>% 
                                      dplyr::filter(cellname == .y) %>% 
                                      unnest(data),
                                    df_co2,
                                    df_tseries_sampled %>% 
                                      dplyr::filter(cellname == .y),
                                    params_siml, 
                                     params_modl, 
                                    df_soiltexture,
                                    agg = FALSE
                                  ))) %>% 
  dplyr::select(-data)

save(df_pmodel_zanitest, file = "data/df_pmodel_zanitest.RData")

df_pmodel_zanitest_flat <- df_pmodel_zanitest %>% 
  unnest(out_pmodel)
write_csv(df_pmodel_zanitest_flat, path = "data/df_pmodel_zanitest_flat.csv")
```

Plot example: one species/site/year.
```{r}
df_pmodel_zanitest_flat %>% 
  dplyr::filter(id_species_site == "1142_Fagus sylvatica" & year(date) == 1970) %>% 
  ggplot(aes(x = date, y = gpp)) +
  geom_line()
```

Compare to LPJ-GUESS from Zani et al.
```{r}
agg_out <- function(df){
  df %>% 
    group_by(year) %>% 
    summarise(gpp = sum(gpp), 
          rd = sum(rd), 
          apar = sum(apar), 
          alpha = mean(alpha, na.rm = TRUE))
}

df_pmodel_zanitest_agg <- df_pmodel_zanitest %>% 
  mutate(out_pmodel = map(out_pmodel, ~agg_out(.)))

save(df_pmodel_zanitest_agg, file = "data/df_pmodel_zanitest_agg.RData")

df_comb <- df_pmodel_zanitest_agg %>% 
  unnest(out_pmodel) %>% 
  right_join(df_zanitest_sub, by = c("id_species_site", "year")) %>% 
  mutate(gpp_net = gpp - rd)

df_comb %>% analyse_modobs2("cA_tot", "gpp_net", type = "heat")
```



