---
title: "Wrap up of re-analysis of Zani et al"
author: "Beni Stocker"
date: "4/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
#library(rbeni)
library(lme4) #lmer
library(MuMIn) #r.squaredGLMM
library(lmerTest) #p-values
library(effects) #plot effects
library(sjPlot) #plot effects
library(patchwork)

## functions to convert lmer effects plots to ggplot objects
ggplot_on <- function(x){
  df <- tibble(upper = x$`scale(on)`$upper[,1],
               lower = x$`scale(on)`$lower[,1],
               off = x$`scale(on)`$fit[,1],
               on = x$`scale(on)`$x[,1])
  gg <- ggplot() + 
    geom_ribbon(data = df, aes(x = on, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df, aes(on, off), col = "royalblue") +
    theme_classic() +
    labs(x = "SOS (DOY)", y = "EOS (DOY)")
  return(gg)
}

ggplot_year <- function(x){
  df <- tibble(upper = x$`scale(year)`$upper[,1],
               lower = x$`scale(year)`$lower[,1],
               off = x$`scale(year)`$fit[,1],
               year = x$`scale(year)`$x[,1])
  gg <- ggplot() + 
    geom_ribbon(data = df, aes(x = year, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df, aes(year, off), col = "royalblue") +
    theme_classic() +
    labs(x = "Year", y = "EOS (DOY)")
  return(gg)
}

ggplot_cA_tot <- function(x){
  df <- tibble(upper = x$`scale(cA_tot)`$upper[,1],
               lower = x$`scale(cA_tot)`$lower[,1],
               off = x$`scale(cA_tot)`$fit[,1],
               cA_tot = x$`scale(cA_tot)`$x[,1])
  gg <- ggplot() + 
    geom_ribbon(data = df, aes(x = cA_tot, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df, aes(cA_tot, off), col = "royalblue") +
    theme_classic() +
    labs(x = expression(paste(italic("A")[net], " (gC m"^-2, " yr"^-1, ")")), y = "EOS (DOY)")
  
  return(gg)
}

ggplot_gpp_net <- function(x){
  df <- tibble(upper = x$`scale(gpp_net)`$upper[,1],
               lower = x$`scale(gpp_net)`$lower[,1],
               off   = x$`scale(gpp_net)`$fit[,1],
               gpp_net = x$`scale(gpp_net)`$x[,1])
  gg <- ggplot() + 
    geom_ribbon(data = df, aes(x = gpp_net, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df, aes(gpp_net, off), col = "royalblue") +
    theme_classic() +
    labs(x = expression(paste(italic("A")[net], " (gC m"^-2, " yr"^-1, ")")), y = "EOS (DOY)")
  
  return(gg)
}

ggplot_year_anet <- function(x, y){
  df <- tibble(upper = x$`scale(year)`$upper[,1],
               lower = x$`scale(year)`$lower[,1],
               off = x$`scale(year)`$fit[,1],
               year = x$`scale(year)`$x[,1]) %>% 
    bind_rows(
      tibble(upper = y$`scale(year)`$upper[,1],
             lower = y$`scale(year)`$lower[,1],
             anet   = y$`scale(year)`$fit[,1],
             year = y$`scale(year)`$x[,1])
    )
  gg <- ggplot() + 
    geom_ribbon(data = df, aes(x = year, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df, aes(year, off), col = "royalblue") +
    theme_classic() +
    labs(x = "Year", y = "EOS (DOY)")
  return(gg)
}
```

## Load data

Read the data complemented with drivers, obtained from Constantin Zohner (7.12.2020).
```{r}
df <- data.table::fread("~/data/pep/processed/DataMeta_3_Drivers_20_11_10.csv") %>% 
  as_tibble() %>% 
  rename(lon = LON, lat = LAT, year = YEAR, off = DoY_off, on = DoY_out, 
         anom_off_zani = autumn_anomaly, anom_on_zani = spring_anomaly, 
         species = Species, id_site = PEP_ID, id_species_site = timeseries)

min(df_orig$year) 
  # ## use the non-water-stressed version of A
  # mutate(cA_tot = `cA_tot-w`)

# ## test
# tmp <- df %>% 
#   dplyr::select(id_site, id_species_site) %>% 
#   group_by(id_site) %>% 
#   nest() %>% 
#   mutate(data = map(data, ~distinct(.)))
```

Taking uncorrected Zani
```{r}
df_anl <- df
```

Correct Zani EOS dates
```{r}
## what does the original data say?
load("~/data/pep/processed/df_PEP725_alldata.RData") # loads 'df'
df_orig <- df %>% 
  dplyr::filter(phase_id %in% c(10, 11, 13, 95, 205)) %>% 
  mutate(pheno = ifelse(phase_id %in% c(10, 11, 13), "on", ifelse(phase_id %in% c(95, 205), "off", NA))) %>% 
  rename(id_site = s_id)

## take the mean across inconsistent duplicates. 3,223,360 -> 2,987,845 data points.
df_orig <- df_orig %>% 
  dplyr::select(id_site, species, year, pheno, day) %>% 
  group_by(id_site, species, year, pheno) %>% 
  summarise(day = round(mean(day), 0))

df_orig_wide <- df_orig %>% 
   pivot_wider(names_from = "pheno", values_from = "day")

df_anl <- df_anl %>% 
  rename(on_zani = on, off_zani = off) %>% 
  left_join(df_orig_wide, by = c("id_site", "species", "year"))

out <- df_anl %>% analyse_modobs2("off", "off_zani", type = "heat")
out$gg
```

Add P-model outputs.
```{r}
load("~/data/pep/processed/df_pmodel_zani.RData")

df_anl <- df_pmodel_zani %>% 
  unnest(out_pmodel) %>% 
  right_join(df_anl, by = c("id_species_site", "year")) %>% 
  mutate(gpp_net = gpp - rd,
         lue = gpp / apar)
```

Check consistency of P-model and LPJ-GUESS GPP.
```{r}
out <- df_anl %>% 
  analyse_modobs2("cA_tot", "gpp_net", type = "heat")
out$gg
```

## Data volume

Distribution of number of data points per time series.
```{r}
df_ndata <- df_anl %>% 
  group_by(id_species_site) %>% 
  summarise(count = n())

quantile(df_ndata$count, probs = c(0.5, 0.9))

df_ndata %>% 
  ggplot(aes(count, ..count..)) +
  geom_histogram(bins = 20) +
  xlab("N data points per time series (years)") +
  labs(title = "PEP725 phenology time series lengthts")
```

```{r}
yrrange <- df_anl %>% 
  pull(year) %>% 
  range()
yrrange[2] - yrrange[1]

df_anl %>% 
  ggplot(aes(year, ..count..)) +
  geom_histogram() +
  xlab("Year") +
  labs(title = "PEP725 phenology time series lengthts")
```

Range of  latitudes.
```{r}
latrange <- df_anl %>% 
  pull(lat) %>% 
  range()
latrange[2] - latrange[1]
```

## IAV

### SOS

```{r}
Fit_IAV_on = lmer(off ~ scale(on) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")

aic_iav_zz <- AIC(Fit_IAV_on)
out_plot <- allEffects(Fit_IAV_on)
gg_iav_sos <- ggplot_on(out_plot)
gg_iav_sos
```

### Anet LPJ

```{r}
Fit_IAV_cA_tot = lmer(off ~ scale(cA_tot) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")

aic_iav_zz <- AIC(Fit_IAV_cA_tot)
out_plot <- allEffects(Fit_IAV_cA_tot)
gg_iav_zz <- ggplot_cA_tot(out_plot)
gg_iav_zz
```

#### Running mean Anet LPJ

```{r}
df_anl <- df_anl %>% 
  mutate(cA_tot_rm2 = data.table::frollmean(cA_tot, 2),
         cA_tot_rm3 = data.table::frollmean(cA_tot, 3),
         cA_tot_rm5 = data.table::frollmean(cA_tot, 5),
         cA_tot_rm10 = data.table::frollmean(cA_tot, 10))

## test
# df_anl %>% 
#   dplyr::filter(id_species_site == "1_Fagus sylvatica") %>% 
#   ggplot() +
#   geom_line(aes(year, cA_tot)) +
#   geom_line(aes(year, cA_tot_rm3), color = "red") +
#   geom_line(aes(year, cA_tot_rm5), color = "blue") +
#   geom_line(aes(year, cA_tot_rm10), color = "green")

fit2 = lmer(off ~ scale(cA_tot_rm2) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
fit3 = lmer(off ~ scale(cA_tot_rm3) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
fit5 = lmer(off ~ scale(cA_tot_rm5) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")
fit10 = lmer(off ~ scale(cA_tot_rm10) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")

out_plot2 <- allEffects(fit2)
out_plot3 <- allEffects(fit3)
out_plot5 <- allEffects(fit5)
out_plot10 <- allEffects(fit10)

ggplot_cA_tot_rm <- function(out_plot2, out_plot3, out_plot5, out_plot10){
    
  df <- tibble(upper = out_plot2$`scale(cA_tot_rm2)`$upper[,1],
               lower = out_plot2$`scale(cA_tot_rm2)`$lower[,1],
               off = out_plot2$`scale(cA_tot_rm2)`$fit[,1],
               cA_tot = out_plot2$`scale(cA_tot_rm2)`$x[,1]) %>% 
    mutate(nyears = 2) %>% 
    bind_rows(
      tibble(upper = out_plot3$`scale(cA_tot_rm3)`$upper[,1],
               lower = out_plot3$`scale(cA_tot_rm3)`$lower[,1],
               off = out_plot3$`scale(cA_tot_rm3)`$fit[,1],
               cA_tot = out_plot3$`scale(cA_tot_rm3)`$x[,1]) %>% 
    mutate(nyears = 3)
    ) %>% 
    bind_rows(
      tibble(upper = out_plot5$`scale(cA_tot_rm5)`$upper[,1],
               lower = out_plot5$`scale(cA_tot_rm5)`$lower[,1],
               off = out_plot5$`scale(cA_tot_rm5)`$fit[,1],
               cA_tot = out_plot5$`scale(cA_tot_rm5)`$x[,1]) %>% 
    mutate(nyears = 5)
    ) %>% 
    bind_rows(
      tibble(upper = out_plot10$`scale(cA_tot_rm10)`$upper[,1],
               lower = out_plot10$`scale(cA_tot_rm10)`$lower[,1],
               off = out_plot10$`scale(cA_tot_rm10)`$fit[,1],
               cA_tot = out_plot10$`scale(cA_tot_rm10)`$x[,1]) %>% 
    mutate(nyears = 10)
    )
  
  gg <- ggplot() + 
    # geom_ribbon(data = df, aes(x = cA_tot, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df, aes(x = cA_tot, y = off, color = nyears, group = nyears)) +
    theme_classic() +
    labs(x = expression(paste(italic("A")[net], " (gC m"^-2, "yr"^-2, ")")), y = "EOS (DOY)")
}

gg_2yr_zz <- ggplot_cA_tot_rm(out_plot2, out_plot3, out_plot5, out_plot10)
gg_2yr_zz
```

### Anet P-model

```{r}
fit = lmer(off ~ scale(gpp_net) + (1|id_site) + (1|species) , data = df_anl, na.action = "na.exclude")

aic_iav_us <- AIC(fit)
out_plot <- allEffects(fit)
gg_iav_us <- ggplot_gpp_net(out_plot)
gg_iav_us
```


## IAV vs. long-term

### SOS

```{r}
Fit_LT_on = lmer(off ~ scale(on) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")

aic_lt_zz <- AIC(Fit_LT_on)
out_plot <- allEffects(Fit_LT_on)
gg_lt_sos_on <- ggplot_on(out_plot)
gg_lt_sos_year <- ggplot_year(out_plot)

gg_lt_sos_on + gg_lt_sos_year
```

Long-term trend in SOS.
```{r}
mod_on_year <- lmer(on ~ scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")

out_plot <- allEffects(mod_on_year)
gg_lt_sos_year_t <- ggplot_year(out_plot)

gg_lt_sos_year_t <- gg_lt_sos_year_t + ylab("SOS (DOY)") + 
  theme(plot.background = element_rect(fill = "grey80"))

gg_lt_sos_year_t
```

### Anet LPJ

```{r}
Fit_LT_cA_tot = lmer(off ~ scale(cA_tot) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")

aic_lt_zz <- AIC(Fit_LT_cA_tot)
out_plot_lt <- allEffects(Fit_LT_cA_tot)
gg_lt_zz_anet <- ggplot_cA_tot(out_plot_lt)
gg_lt_zz_year <- ggplot_year(out_plot_lt)

gg_lt_zz_anet + gg_lt_zz_year
```
Is this the better model than EOS ~ Anet?
```{r}
out_anova <- anova(Fit_IAV_cA_tot, Fit_LT_cA_tot)
out_anova
```


Long-term trend in Anet
```{r}
mod_anet_year <- lmer(cA_tot ~ scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")

out_plot_anet <- allEffects(mod_anet_year)
gg_lt_anet_year <- ggplot_year(out_plot_anet)

gg_lt_anet_year <- gg_lt_anet_year + 
  labs(y = expression(italic("A")[net])) + 
  theme(plot.background = element_rect(fill = "grey80"))
gg_lt_anet_year
```


```{r}
ggplot_year_anet <- function(x, y){
  
  df1 <- tibble(upper = x$`scale(year)`$upper[,1],
               lower = x$`scale(year)`$lower[,1],
               off = x$`scale(year)`$fit[,1],
               year = x$`scale(year)`$x[,1])
  df2 <- tibble(upper = y$`scale(year)`$upper[,1],
             lower = y$`scale(year)`$lower[,1],
             anet = y$`scale(year)`$fit[,1],
             year = y$`scale(year)`$x[,1])
  
  coef <- 0.7 * mean(df1$off) / mean(df2$anet)
  
  # library(grid)
  # t1 <- textGrob(expression(paste("EOS ~ ", italic("A")[net], " + ", bold("Year"))), x = 0, y = 1.1, gp = gpar(col = "black"))
  
  gg <- ggplot() + 
    geom_ribbon(data = df1, aes(x = year, ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(data = df1, aes(x = year, y = off), col = "royalblue") +
    
    geom_ribbon(data = df2, aes(x = year, ymin = lower * coef + 100, ymax = upper * coef + 100), alpha = 0.2) +
    geom_line(data = df2, aes(x = year, y = anet * coef + 100), col = "red") +
    scale_y_continuous(name = "EOS (DOY)",
                       sec.axis = sec_axis( ~(. - 100) / coef, name = expression(paste(italic("A")[net], " (gC m"^-2, " yr"^-1, ")")))) +
    labs(x = "Year", 
         title = expression(paste("EOS ~ ", italic("A")[net], " + ", bold("Year"), "      ", italic("A")[net], " ~ Year")), 
         subtitle = "Zani et al.") +
    theme_classic() +
    # annotation_custom(grobTree(t1)) +
    theme(axis.text.y.right = element_text(colour="red"),
          axis.ticks.y.right = element_line(colour="red"),
          axis.title.y.right = element_text(colour="red"),
          axis.text.y = element_text(colour="royalblue"),
          axis.ticks.y = element_line(colour="royalblue"),
          axis.title.y = element_text(colour="royalblue"),
          axis.line.y = element_line(colour = "royalblue"),
          axis.line.y.right = element_line(colour = "red"))
  
  return(gg)
}

gg_dual <- ggplot_year_anet(x = out_plot_lt, y = out_plot_anet)
```

### Anet P-model

```{r}
load("data/df_anl_pmodel.RData")
fit = lmer(off ~ scale(gpp_net) + scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")

summary(fit)
aic_lt_zz <- AIC(fit)
out_plot <- allEffects(fit)
gg_lt_us_anet <- ggplot_gpp_net(out_plot)
gg_lt_us_year <- ggplot_year(out_plot)

gg_lt_us_anet + gg_lt_us_year
```

Long-term trend in Anet
```{r}
mod_anet_year <- lmer(gpp_net ~ scale(year) + (1|id_site) + (1|species), data = df_anl, na.action = "na.exclude")

out_plot <- allEffects(mod_anet_year)
gg_lt_anet_year <- ggplot_year(out_plot)

gg_lt_anet_year <- gg_lt_anet_year + 
  labs(y = expression(italic("A")[net])) + 
  theme(plot.background = element_rect(fill = "grey80"))
```

## IAV vs. long-term vs. mean

This is based on a spatially stratified sampling.

### Anet LPJ

Separate mean across years 1970-2000 from interannual anomaly.
```{r eval=FALSE}
separate_anom <- function(df){
  df_mean <- df %>% 
    dplyr::filter(year %in% 1970:2000) %>% 
    summarise(mean_off = mean(off, na.rm = TRUE), 
              mean_cA_tot = mean(cA_tot, na.rm = TRUE))
  df %>% 
    mutate(mean_cA_tot = df_mean$mean_cA_tot,
           anom_cA_tot = cA_tot - df_mean$mean_cA_tot)
}
df_anl <- df_anl %>% 
  group_by(id_species_site) %>% 
  nest() %>% 
  mutate(data = map(data, ~separate_anom(.))) %>% 
  unnest(data)
```

Bootstrap 300 samples and test for significance in the mean Anet coefficient. This returns a data frame containing the coefficient estimate and the p-value for each sample.
```{r eval=FALSE}
get_pval_mean_anet <- function(use_seed, df){
  
  ## sample one time series per gridcell
  set.seed(use_seed)
  df_sites_sampled <- df %>% 
    dplyr::select(id_species_site, lon_mid, lat_mid) %>% 
    distinct() %>% 
    group_by(lon_mid, lat_mid) %>% 
    sample_n(1)
  df <- df %>% 
    dplyr::filter(id_species_site %in% df_sites_sampled$id_species_site)
  
  ## fit model separating mean Anet
  Fit_LT_anom = lmer(off ~ scale(mean_cA_tot) + scale(anom_cA_tot) + scale(year) + (1|id_site) + (1|species), data = df, na.action = "na.exclude")
  
  ## return p value for significance of term representing mean Anet
  out_sum <- summary(Fit_LT_anom)
  pval <- out_sum$coefficients["scale(mean_cA_tot)", "Pr(>|t|)"]
  esti <- out_sum$coefficients["scale(mean_cA_tot)", "Estimate"]
  
  return(tibble(estimate = esti, pval = pval))
}

n_samples <- 300
df_sampled_mean_anet <- purrr::map_dfr(as.list(seq(n_samples)), ~get_pval_mean_anet(., df_anl))

df_sampled_mean_anet %>% 
  ggplot(aes(estimate, ..count..)) +
  geom_histogram()

n_sign <- df_sampled_mean_anet %>% 
  dplyr::filter(pval < 0.05) %>% 
  nrow()

## fraction not significant
1 - n_sign/n_samples
```


## Figure 1

```{r}
gg_iav_zz <- gg_iav_zz +
  labs(title = expression(paste("EOS ~ ", italic("A")[net])), subtitle = "Zani et al.")

gg_iav_us <- gg_iav_us +
  labs(title = expression(paste("EOS ~ ", italic("A")[net])), subtitle = "P-model")

gg_lt_zz_anet <- gg_lt_zz_anet +
  labs(title = expression(paste("EOS ~ ", bolditalic("A")[bold(net)], " + Year")), subtitle = "Zani et al.")

gg_lt_zz_year <- gg_lt_zz_year +
  labs(title = expression(paste("EOS ~ ", italic("A")[net], " + ", bold("Year"))), subtitle = "Zani et al.")

pp <- (gg_iav_zz + gg_iav_us) /
  (gg_lt_zz_anet + gg_dual)
pp + plot_annotation(tag_levels = 'A')
ggsave("fig/fig_1_wrapup_zz.pdf", width = 8, height = 6)
ggsave("fig/fig_1_wrapup_zz.png", width = 8, height = 6)

pp <- (gg_iav_sos + gg_lt_sos_year_t) /
  (gg_lt_sos_on + gg_lt_sos_year)
pp + plot_annotation(
  title = expression(paste("EOS ~ SOS")),
  tag_levels = 'A'
)
ggsave("fig/fig_1_wrapup_sos.pdf", width = 8, height = 6)
ggsave("fig/fig_1_wrapup_sos.png", width = 8, height = 6)

pp <- (gg_iav_us + gg_lt_anet_year) /
  (gg_lt_us_anet + gg_lt_us_year)
pp + plot_annotation(
  title = expression(paste("EOS ~ ", italic("A")[net], " P-model")),
  tag_levels = 'A'
)
ggsave("fig/fig_1_wrapup_sos.pdf", width = 8, height = 6)
ggsave("fig/fig_1_wrapup_sos.png", width = 8, height = 6)
```


Fig. 1 (A) Relationship of end-of-season (EOS, expressed as day-of-year, DOY) versus Anet simulated by Zani et al., based on a linear mixed-effects regression (LMER) model with scaled Anet as a fixed effect and site and species as grouping variables of random intercepts. (B) Temporal trend of Anet, based on a LMER model with year as a fixed effect and site and species as grouping variables of the random intercept. (C, D) Partial relationships of a multiple LMER model, where the long-term trend (year) and Anet are treated as fixed effects (effects shown here), and site and species are treated as grouping variables of the random intercepts. The trend towards later EOS (D) exists in parallel with the trend towards higher Anet (B).
