---
title: "Compare predictions"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# Load libraries
library(data.table)
library(tidyverse)
library(broom)
library(Metrics)
library(lmodel2)
library(ggplot2)
if(!require(devtools)){install.packages(devtools)} 
devtools::install_github("bluegreen-labs/phenor@v1.0")
library(phenor)
#detach(package:phenor)
library(ggpubr)
```

This script compares the predicted autumn senescense read from the ModelAnalysis_1_Predicted_DoYoff_subset.csv file provided by Zani et al. in Zenodo with the predictions resulting from evaluating the models using the optimal parameters read from ModelAnalysis_2_OptimalParameters_subset.csv and data.sub for 3 timeseries.

### A. Load subset of data for the demo

```{r,include=FALSE,eval=TRUE}
load("~/pep/calibration/data/datalist_demo.RData")
load("~/pep/calibration/data/preds_df_demo.RData")
load("~/pep/calibration/data/df_tseries_id_demo.RData")
```

### B. Define the models

The models are defined as in Zani et al. 2020. For this demo, only the TPM, TPDM and PIA+ models are considered.

```{r, include=FALSE,eval=TRUE}

## MODELS OF LEAF SENESCENCE (and drivers):

## First-generation:
# CDD (chilling temperature) - Dufrene et al. (2005)
# DM1 and DM2 (chilling temperature, autumn daylength) - Delpierre et al. (2009)
# TPM (chilling temperature, autumn daylength) - Lang et al. (2019)
## Second-generation:
# SIAM (chilling temperature, autumn daylength, spring anomaly) - Keenan and Richardson (2015)
# TDM and TPDM (chilling temperature, autumn daylength, growing season temperature / + water stress) - Liu et al. (2019)
## PIA:
# PIA_gsi (chilling temperature, autumn daylength, leaf flushing date, growing season mean temperature, daylength, vapour pressure deficit)
# PIA-/+ (chilling temperature, autumn daylength, leaf flushing date, growing season mean temperature, daylength, precipitation, net radiation, CO2 concentration, -/+ water stress)

TPM.model = function(par, data){
  # exit the routine as some parameters are missing
  if (length(par) != 4){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the par argument for readability
  P_base = par[1]
  a = par[2]
  b = par[3]
  F_crit = par[4]
  
  # create forcing/chilling rate vector at the day level
  Rf = 1/(1+exp(a*(data$Tmini*data$Li-b)))
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when photoperiod is shorter than the photoperiod threshold (P_base)
  # after the date of the longest photoperiod (summer solstice), namely, the 173rd day of year
  t0 <- vector()
  for(c in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Li[,c] < P_base)]
    ind1 = min(which(t0A > 173))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  
  # nullify values before the t0
  for(c in 1:ncol(data$Tmini)){
    Rf[1:t0[c],c] = 0 #nullify values before the date of leaf.out
  }
  
  # calculate the summation along the year and derive the date of leaf.off
  # DOY of budburst criterium
  doy = apply(Rf,2, function(xt){
    doy = which(cumsum(xt) >= F_crit)[1]
  })
  
  return(doy)
}

SecondGen_PIA.models = function(par, predictor, data) {
  # exit the routine as some parameters are missing
  if (length(par) != 5 & length(par) != 6){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the
  # par argument for readability
  P_base = as.numeric(par[1])
  a = as.numeric(par[2])
  b = as.numeric(par[3])
  c = as.numeric(par[4])
  if(length(par)==5) {
    d = as.numeric(par[5])
    pred = predictor
  }
  if(length(par)==6) {
    d = as.numeric(par[5])
    e = as.numeric(par[6])
    pred1 = predictor[[1]] #xxx? predictor[[1]] predictor is given as a vector in the calibration and predictions and should be a list
    pred2 = predictor[[2]] #xxx? predictor[[2]] predictor is given as a vector in the calibration and predictions and should be a list
  }
  
  # create forcing/chilling rate vector at the day level
  Rf = 1/(1+exp(a*(data$Tmini*data$Li-b)))
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when photoperiod is shorter than the photoperiod threshold (P_base)
  # after the date of the longest photoperiod (summer solstice), namely, the 173rd day of year
  t0 <- vector()
  for(col in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Li[,col] < P_base)] #xxx? P_base instead of 173
    ind1 = min(which(t0A > 173))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  
  # nullify values before the t0
  for(col in 1:ncol(data$Tmini)){
    Rf[1:t0[col],col] = 0 
  }
  
  if(length(par)==5) {
    # add predictor at the end of the matrix-columns
    Rf = rbind(Rf,pred) # pred instead of predictor
    
    # predict date of leaf.off
    doy = apply(Rf,2, function(xt){
      doy = which(cumsum(xt[1:366]) >= c+d*xt[367])[1]
    }) 
  }
  if(length(par)==6) {
    # add predictors at the end of the matrix-columns
    Rf = rbind(Rf,pred1)
    Rf = rbind(Rf,pred2)
    
    # predict date of leaf.off
    doy = apply(Rf,2, function(xt){
      doy = which(cumsum(xt[1:366]) >= c+d*xt[367]+e*xt[368])[1]
    }) 
  }
  
  return(doy)
}

```

### C. Example 1. Timeseries: 1_Fagus_sylvatica

1. Select data for the timeseries

```{r ,include=FALSE,eval=TRUE}
args = commandArgs(trailingOnly=TRUE)

##------------------------------------------
## Select a timeseries
##------------------------------------------
args <- as.character(df_tseries_id$idx[which(df_tseries_id$timeseries == "1_Fagus_sylvatica")])

##------------------------------------------
## given timeseries number, determine species and pep_id given args[1]
##------------------------------------------
df_tmp <- df_tseries_id %>% 
  dplyr::filter(idx == as.integer(args[1]))
  
sp <- df_tmp %>% pull(Species)
site <- df_tmp %>% pull(PEP_ID)

##------------------------------------------
## Subset according to timeseries
##------------------------------------------
ts <- paste0(site,"_",sp) 
data.sub <- DataList[[sp]][[as.character(site)]]
data.sub$Tmini <- unname(data.sub$Tmini, force = FALSE)
data.sub$Li <- unname(data.sub$Li, force = FALSE)
preds.sub <- preds.df %>% dplyr::filter(timeseries==ts)
```

2. Plot observed versus predicted autumn senescence dates for the timeseries.

a) The predictions are read from the ModelAnalysis_1_Predicted_DoYoff_subset.csv file provided by Zani et al. in Zenodo. 
   These predictions are calculated using the function estimate_phenology from the phenor package.

```{r, include=FALSE, eval=TRUE}

# Reading sample outputs from Zani et al. for predictions
pred_DoYoff <- data.table::fread("~/pep/calibration/output/ModelAnalysis_1_Predicted_DoYoff_subset.csv") 
subsetPred_1 <- pred_DoYoff %>% filter(timeseries=="1_Fagus sylvatica")

PlotTPM <- ggplot(data = subsetPred_1, aes(x = Pred_AnomDoYoff_TPM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "Predicted from the phenor pkg function",
              subtitle = "Model TPM",
              caption = "")

PlotTPDM <- ggplot(data = subsetPred_1, aes(x = Pred_AnomDoYoff_TPDM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly ") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

PlotPIAplus <- ggplot(data = subsetPred_1, aes(x = `Pred_AnomDoYoff_PIA+`, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+",
              caption = "Data source: Predictions from ModelAnalysis_1_Predicted_DoYoff_subset.csv")
  
```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

b) The predictions are calculated from the optimal parameters read from the ModelAnalysis_2_OptimalParameters_subset.csv file provided by Zani et al. in Zenodo. 
   These predictions are the result of evaluating the models specified in the paper for the optimal parameters and the data.

```{r,include=FALSE, eval=TRUE}

# Reading sample outputs from Zani et al. for predictions
opt_param <- data.table::fread("~/pep/calibration/output/ModelAnalysis_2_OptimalParameters_subset.csv") 
subset_opt_param_1 <- opt_param %>% filter(timeseries=="1_Fagus sylvatica")

# Gathering the parameters for TPM together
parTPM <- c(subset_opt_param_1$Pbase_TPM,subset_opt_param_1$a_TPM,subset_opt_param_1$b_TPM,subset_opt_param_1$Fcrit_TPM)
# Evaluating the TPM model with the optimal parameters and the data for this timeseries
predfromparam.TPM.model = TPM.model(par = parTPM, data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_1$predfromparam.TPM.model <- predfromparam.TPM.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_1$predfromparam.AnomDoYoff_TPM <- subsetPred_1$predfromparam.TPM.model - subsetPred_1$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotTPM <- ggplot(data = subsetPred_1, aes(x = predfromparam.AnomDoYoff_TPM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "Predicted from model evaluation",
              subtitle = "Model TPM",
              caption = "")

# Gathering the parameters for TPDM together
parTPDM <- c(subset_opt_param_1$Pbase_TPDM,subset_opt_param_1$a_TPDM,subset_opt_param_1$b_TPDM,subset_opt_param_1$c_TPDM,subset_opt_param_1$d_TPDM,
             subset_opt_param_1$e_TPDM)
# Evaluating the TPDM model with the optimal parameters and the data for this timeseries
predfromparam.TPDM.model = SecondGen_PIA.models(par = parTPDM, predictor = list(preds.sub$temp_GS,preds.sub$RD_summer), data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_1$predfromparam.TPDM.model <- predfromparam.TPDM.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_1$predfromparam.AnomDoYoff_TPDM <- subsetPred_1$predfromparam.TPDM.model - subsetPred_1$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotTPDM <- ggplot(data = subsetPred_1, aes(x = predfromparam.AnomDoYoff_TPDM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

# Gathering the parameters for PIA+ together
parPIAplus <- c(subset_opt_param_1$`Pbase_PIA+`,subset_opt_param_1$`a_PIA+`,subset_opt_param_1$`b_PIA+`,subset_opt_param_1$`c_PIA+`,subset_opt_param_1$`d_PIA+`)
# Evaluating the PIA+ model with the optimal parameters and the data for this timeseries
predfromparam.PIAplus.model = SecondGen_PIA.models(par = parPIAplus, predictor = preds.sub$cA_tot, data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_1$predfromparam.PIAplus.model <- predfromparam.PIAplus.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_1$predfromparam.AnomDoYoff_PIAplus <- subsetPred_1$predfromparam.PIAplus.model - subsetPred_1$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotPIAplus <- ggplot(data = subsetPred_1, aes(x = predfromparam.AnomDoYoff_PIAplus, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+",
              caption = "Data source: Predictions from optim params (ModelAnalysis_2_OptimalParameters_subset.csv) and data.sub evaluated in the models")

```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

c) These figures plot both predictions calculated using the phenor pck functions vs. predictions calculated from the optimal parameters.

```{r,include=FALSE,eval=TRUE}
# Plotting observations vs. predictions from parameters
PlotTPM <- ggplot(data = subsetPred_1, aes(x = predfromparam.AnomDoYoff_TPM, y = Pred_AnomDoYoff_TPM)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "Predicted from model evaluation",
              subtitle = "Model TPM",
              caption = "")

PlotTPDM <- ggplot(data = subsetPred_1, aes(x = predfromparam.AnomDoYoff_TPDM, y = Pred_AnomDoYoff_TPDM)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

PlotPIAplus <- ggplot(data = subsetPred_1, aes(x = predfromparam.AnomDoYoff_PIAplus, y = `Pred_AnomDoYoff_PIA+`)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+")
```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

### Example 2. Timeseries: 1_Quercus_robur

1. Select data for the timeseries

```{r ,include=FALSE,eval=TRUE}
args = commandArgs(trailingOnly=TRUE)

##------------------------------------------
## Select a timeseries
##------------------------------------------
args <- as.character(df_tseries_id$idx[which(df_tseries_id$timeseries == "1_Quercus_robur")])

##------------------------------------------
## given timeseries number, determine species and pep_id given args[1]
##------------------------------------------
df_tmp <- df_tseries_id %>% 
  dplyr::filter(idx == as.integer(args[1]))
  
sp <- df_tmp %>% pull(Species)
site <- df_tmp %>% pull(PEP_ID)

##------------------------------------------
## Subset according to timeseries
##------------------------------------------
ts <- paste0(site,"_",sp) 
data.sub <- DataList[[sp]][[as.character(site)]]
data.sub$Tmini <- unname(data.sub$Tmini, force = FALSE)
data.sub$Li <- unname(data.sub$Li, force = FALSE)
preds.sub <- preds.df %>% dplyr::filter(timeseries==ts)
```

2. Plot observed versus predicted autumn senescence dates for the timeseries.

a) The predictions are read from the ModelAnalysis_1_Predicted_DoYoff_subset.csv file provided by Zani et al. in Zenodo. 
   These predictions are calculated using the function estimate_phenology from the phenor package.

```{r, include=FALSE, eval=TRUE}

# Reading sample outputs from Zani et al. for predictions
pred_DoYoff <- data.table::fread("~/pep/calibration/output/ModelAnalysis_1_Predicted_DoYoff_subset.csv") 
subsetPred_2 <- pred_DoYoff %>% filter(timeseries=="1_Quercus robur")

PlotTPM <- ggplot(data = subsetPred_2, aes(x = Pred_AnomDoYoff_TPM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "Predicted from the phenor pkg function",
              subtitle = "Model TPM",
              caption = "")


PlotTPDM <- ggplot(data = subsetPred_2, aes(x = Pred_AnomDoYoff_TPDM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly ") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

PlotPIAplus <- ggplot(data = subsetPred_2, aes(x = `Pred_AnomDoYoff_PIA+`, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+",
              caption = "Data source: Predictions from ModelAnalysis_1_Predicted_DoYoff_subset.csv")

```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

b) The predictions are calculated from the optimal parameters read from the ModelAnalysis_2_OptimalParameters_subset.csv file provided by Zani et al. in Zenodo. 
   These predictions are the result of evaluating the models specified in the paper for the optimal parameters and the data.

```{r,include=FALSE, eval=TRUE}

# Reading sample outputs from Zani et al. for predictions
opt_param <- data.table::fread("~/pep/calibration/output/ModelAnalysis_2_OptimalParameters_subset.csv") 
subset_opt_param_2 <- opt_param %>% filter(timeseries=="1_Quercus robur")

# Gathering the parameters for TPM together
parTPM <- c(subset_opt_param_2$Pbase_TPM,subset_opt_param_2$a_TPM,subset_opt_param_2$b_TPM,subset_opt_param_2$Fcrit_TPM)
# Evaluating the TPM model with the optimal parameters and the data for this timeseries
predfromparam.TPM.model = TPM.model(par = parTPM, data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_2$predfromparam.TPM.model <- predfromparam.TPM.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_2$predfromparam.AnomDoYoff_TPM <- subsetPred_2$predfromparam.TPM.model - subsetPred_2$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotTPM <- ggplot(data = subsetPred_2, aes(x = predfromparam.AnomDoYoff_TPM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "Predicted from model evaluation",
              subtitle = "Model TPM",
              caption = "")

# Gathering the parameters for TPDM together
parTPDM <- c(subset_opt_param_2$Pbase_TPDM,subset_opt_param_2$a_TPDM,subset_opt_param_2$b_TPDM,subset_opt_param_2$c_TPDM,subset_opt_param_2$d_TPDM,
             subset_opt_param_2$e_TPDM)
# Evaluating the TPDM model with the optimal parameters and the data for this timeseries
predfromparam.TPDM.model = SecondGen_PIA.models(par = parTPDM, predictor = list(preds.sub$temp_GS,preds.sub$RD_summer), data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_2$predfromparam.TPDM.model <- predfromparam.TPDM.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_2$predfromparam.AnomDoYoff_TPDM <- subsetPred_2$predfromparam.TPDM.model - subsetPred_2$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotTPDM <- ggplot(data = subsetPred_2, aes(x = predfromparam.AnomDoYoff_TPDM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

# Gathering the parameters for PIA+ together
parPIAplus <- c(subset_opt_param_2$`Pbase_PIA+`,subset_opt_param_2$`a_PIA+`,subset_opt_param_2$`b_PIA+`,subset_opt_param_2$`c_PIA+`,subset_opt_param_2$`d_PIA+`)
# Evaluating the PIA+ model with the optimal parameters and the data for this timeseries
predfromparam.PIAplus.model = SecondGen_PIA.models(par = parPIAplus, predictor = preds.sub$cA_tot, data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_2$predfromparam.PIAplus.model <- predfromparam.PIAplus.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_2$predfromparam.AnomDoYoff_PIAplus <- subsetPred_2$predfromparam.PIAplus.model - subsetPred_2$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotPIAplus <- ggplot(data = subsetPred_2, aes(x = predfromparam.AnomDoYoff_PIAplus, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+",
              caption = "Data source: Predictions from optim params (ModelAnalysis_2_OptimalParameters_subset.csv) and data.sub evaluated in the models")

```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

c) These figures plot both predictions calculated using the phenor pck functions vs. predictions calculated from the optimal parameters.

```{r,include=FALSE,eval=TRUE}
# Plotting observations vs. predictions from parameters
PlotTPM <- ggplot(data = subsetPred_2, aes(x = predfromparam.AnomDoYoff_TPM, y = Pred_AnomDoYoff_TPM)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "Predicted from model evaluation",
              subtitle = "Model TPM",
              caption = "")

PlotTPDM <- ggplot(data = subsetPred_2, aes(x = predfromparam.AnomDoYoff_TPDM, y = Pred_AnomDoYoff_TPDM)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

PlotPIAplus <- ggplot(data = subsetPred_2, aes(x = predfromparam.AnomDoYoff_PIAplus, y = `Pred_AnomDoYoff_PIA+`)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+")
```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

### Example 3. Timeseries: 1002_Aesculus_hippocastanum

1. Select data for the timeseries

```{r ,include=FALSE,eval=TRUE}
args = commandArgs(trailingOnly=TRUE)

##------------------------------------------
## Select a timeseries
##------------------------------------------
args <- as.character(df_tseries_id$idx[which(df_tseries_id$timeseries == "1002_Aesculus_hippocastanum")])

##------------------------------------------
## given timeseries number, determine species and pep_id given args[1]
##------------------------------------------
df_tmp <- df_tseries_id %>% 
  dplyr::filter(idx == as.integer(args[1]))
  
sp <- df_tmp %>% pull(Species)
site <- df_tmp %>% pull(PEP_ID)

##------------------------------------------
## Subset according to timeseries
##------------------------------------------
ts <- paste0(site,"_",sp) 
data.sub <- DataList[[sp]][[as.character(site)]]
data.sub$Tmini <- unname(data.sub$Tmini, force = FALSE)
data.sub$Li <- unname(data.sub$Li, force = FALSE)
preds.sub <- preds.df %>% dplyr::filter(timeseries==ts)
```

2. Plot observed versus predicted autumn senescence dates for the timeseries.

a) The predictions are read from the ModelAnalysis_1_Predicted_DoYoff_subset.csv file provided by Zani et al. in Zenodo. 
   These predictions are calculated using the function estimate_phenology from the phenor package.

```{r, include=FALSE, eval=TRUE}

# Reading sample outputs from Zani et al. for predictions
pred_DoYoff <- data.table::fread("~/pep/calibration/output/ModelAnalysis_1_Predicted_DoYoff_subset.csv") 
subsetPred_3 <- pred_DoYoff %>% filter(timeseries=="1002_Aesculus hippocastanum")

PlotTPM <- ggplot(data = subsetPred_3, aes(x = Pred_AnomDoYoff_TPM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "Predicted from the phenor pkg function",
              subtitle = "Model TPM",
              caption = "")


PlotTPDM <- ggplot(data = subsetPred_3, aes(x = Pred_AnomDoYoff_TPDM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly ") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

PlotPIAplus <- ggplot(data = subsetPred_3, aes(x = `Pred_AnomDoYoff_PIA+`, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from phenor pkg fc") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+",
              caption = "Data source: Predictions from ModelAnalysis_1_Predicted_DoYoff_subset.csv")

```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

b) The predictions are calculated from the optimal parameters read from the ModelAnalysis_2_OptimalParameters_subset.csv file provided by Zani et al. in Zenodo. 
   These predictions are the result of evaluating the models specified in the paper for the optimal parameters and the data.

```{r,include=FALSE, eval=TRUE}

# Reading sample outputs from Zani et al. for predictions
opt_param <- data.table::fread("~/pep/calibration/output/ModelAnalysis_2_OptimalParameters_subset.csv") 
subset_opt_param_3 <- opt_param %>% filter(timeseries=="1002_Aesculus hippocastanum")

# Gathering the parameters for TPM together
parTPM <- c(subset_opt_param_3$Pbase_TPM,subset_opt_param_3$a_TPM,subset_opt_param_3$b_TPM,subset_opt_param_3$Fcrit_TPM)
# Evaluating the TPM model with the optimal parameters and the data for this timeseries
predfromparam.TPM.model = TPM.model(par = parTPM, data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_3$predfromparam.TPM.model <- predfromparam.TPM.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_3$predfromparam.AnomDoYoff_TPM <- subsetPred_3$predfromparam.TPM.model - subsetPred_3$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotTPM <- ggplot(data = subsetPred_3, aes(x = predfromparam.AnomDoYoff_TPM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "Predicted from model evaluation",
              subtitle = "Model TPM",
              caption = "")

# Gathering the parameters for TPDM together
parTPDM <- c(subset_opt_param_3$Pbase_TPDM,subset_opt_param_3$a_TPDM,subset_opt_param_3$b_TPDM,subset_opt_param_3$c_TPDM,subset_opt_param_3$d_TPDM,
             subset_opt_param_3$e_TPDM)
# Evaluating the TPDM model with the optimal parameters and the data for this timeseries
predfromparam.TPDM.model = SecondGen_PIA.models(par = parTPDM, predictor = list(preds.sub$temp_GS,preds.sub$RD_summer), data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_3$predfromparam.TPDM.model <- predfromparam.TPDM.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_3$predfromparam.AnomDoYoff_TPDM <- subsetPred_3$predfromparam.TPDM.model - subsetPred_3$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotTPDM <- ggplot(data = subsetPred_3, aes(x = predfromparam.AnomDoYoff_TPDM, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

# Gathering the parameters for PIA+ together
parPIAplus <- c(subset_opt_param_3$`Pbase_PIA+`,subset_opt_param_3$`a_PIA+`,subset_opt_param_3$`b_PIA+`,subset_opt_param_3$`c_PIA+`,subset_opt_param_3$`d_PIA+`)
# Evaluating the PIA+ model with the optimal parameters and the data for this timeseries
predfromparam.PIAplus.model = SecondGen_PIA.models(par = parPIAplus, predictor = preds.sub$cA_tot, data = data.sub)
# Adding this predictions (from model evaluation) to the dataset read in ModelAnalysis_1_Predicted_DoYoff_subset.csv for this timeseries
subsetPred_3$predfromparam.PIAplus.model <- predfromparam.PIAplus.model
# Calculate anomalies subtracting the meansite_DoYoff
subsetPred_3$predfromparam.AnomDoYoff_PIAplus <- subsetPred_3$predfromparam.PIAplus.model - subsetPred_3$meansite_DoYoff
# Plotting observations vs. predictions from parameters
PlotPIAplus <- ggplot(data = subsetPred_3, aes(x = predfromparam.AnomDoYoff_PIAplus, y = Obs_AnomDoYoff)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Observed autumn anomaly") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+",
              caption = "Data source: Predictions from optim params (ModelAnalysis_2_OptimalParameters_subset.csv) and data.sub evaluated in the models")

```

```{r,echo=FALSE,warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```

c) These figures plot both predictions calculated using the phenor pck functions vs. predictions calculated from the optimal parameters.

```{r,include=FALSE,eval=TRUE}
# Plotting observations vs. predictions from parameters
PlotTPM <- ggplot(data = subsetPred_3, aes(x = predfromparam.AnomDoYoff_TPM, y = Pred_AnomDoYoff_TPM)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "Predicted from model evaluation",
              subtitle = "Model TPM",
              caption = "")

PlotTPDM <- ggplot(data = subsetPred_3, aes(x = predfromparam.AnomDoYoff_TPDM, y = Pred_AnomDoYoff_TPDM)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "",
              subtitle = "Model TPDM",
              caption = "")

PlotPIAplus <- ggplot(data = subsetPred_3, aes(x = predfromparam.AnomDoYoff_PIAplus, y = `Pred_AnomDoYoff_PIA+`)) + geom_point() + 
  scale_x_continuous("Predicted from optimal parameters evaluation") + scale_y_continuous("Predicted from phenor pkg") + theme_classic() +
  labs(title = "",
              subtitle = "Model PIA+")
```

```{r,echo=FALSE, warning = FALSE, fig.height=4,fig.width=12, eval=TRUE}
ggarrange(PlotTPM,PlotTPDM,PlotPIAplus, ncol = 3, nrow = 1)
```
