---
title: "predictions_spatial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ingestr)
library(ggplot2)
library(phenor)
library(geosphere)
do_eval <- TRUE
```

Read the data complemented with drivers, obtained from Constantin Zohner (7.12.2020).
```{r}
df <- data.table::fread("~/data/pep/processed/DataMeta_3_Drivers_20_11_10.csv") %>% 
  as_tibble() %>% 
  rename(lon = LON, lat = LAT, year = YEAR, off = DoY_off, on = DoY_out, 
         anom_off_zani = autumn_anomaly, anom_on_zani = spring_anomaly, 
         species = Species, id_site = PEP_ID, id_species_site = timeseries) #%>% 
  
  ## use the on-water-stressed version of A
  #mutate(cA_tot = `cA_tot-w`)
```

Bin the data into 0.5 deg grid points.
```{r}
## bin
dlon <- 1
dlat <- 1
lon_breaks <- seq(from = floor(min(df$lon)), to = ceiling(max(df$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df$lat)), to = ceiling(max(df$lat)), by = dlat)

df <- df %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2)

#View(select(df, lon, lat, ilon, ilat, lon_lower, lon_upper, lat_lower, lat_upper, lon_mid, lat_mid))

df <- df %>% 
  group_by(lon_mid, lat_mid) %>% 
  drop_na(lon_mid, lat_mid)
```

Sample one site per gridcell.
```{r}
set.seed(189)
df_sites_sampled <- df %>% 
  dplyr::select(id_site, lon_mid, lat_mid) %>% 
  distinct() %>% 
  group_by(lon_mid, lat_mid) %>% 
  sample_n(1)
```

Subset original data frame based on this selection.
```{r}
df_sampled <- df %>% dplyr::filter(id_site %in% df_sites_sampled$id_site)
```

## MODELS OF LEAF SENESCENCE (and drivers):
This code is as in Zani et al. 2020
```{r}
# CDD (chilling temperature) - Dufrene et al. (2005)
# DM1 and DM2 (chilling temperature, autumn daylength) - Delpierre et al. (2009)
# TPM (chilling temperature, autumn daylength) - Lang et al. (2019)
## Second-generation:
# SIAM (chilling temperature, autumn daylength, spring anomaly) - Keenan and Richardson (2015)
# TDM and TPDM (chilling temperature, autumn daylength, growing season temperature / + water stress) - Liu et al. (2019)
## PIA:
# PIA_gsi (chilling temperature, autumn daylength, leaf flushing date, growing season mean temperature, daylength, vapour pressure deficit)
# PIA-/+ (chilling temperature, autumn daylength, leaf flushing date, growing season mean temperature, daylength, precipitation, net radiation, CO2 concentration, -/+ water stress)

# Define functions of Autumn phenology Models
# Modified from https://github.com/khufkens/phenor/blob/master/R/phenology_models.R
CDD.model = function(par, data){
  # exit the routine as some parameters are missing
  if (length(par) != 2){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the
  # par argument for readability
  T_base = par[1]
  F_crit = par[2]
  
  # create forcing/chilling rate vector
  Rf = data$Tmini - T_base
  Rf[Rf > 0] = 0
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when daily minimum temperature is lower than a temperature threshold (T_base) 
  # after the date of the peak multiyear average daily minimum temperature, namely the 200th day of year
  t0 <- vector()
  for(c in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Tmini[,c] < T_base)]
    ind1 = min(which(t0A > 200))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  # nullify values before the t0
  for(c in 1:ncol(data$Tmini)){
    Rf[1:t0[c],c] = 0 
  }
  
  # predict date of leaf.off according to optimized F_crit
  doy = apply(Rf,2, function(xt){
    doy = which(cumsum(xt) <= F_crit)[1]
  })
  
  return(doy)
}
DM1.model = function(par, data){
  # exit the routine as some parameters are missing
  if (length(par) != 3){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the
  # par argument for readability
  T_base = par[1]
  P_base = par[2]
  F_crit = par[3]
  
  # create forcing/chilling rate vector at the day level
  Rf = (data$Tmini - T_base)*(data$Li/P_base) # lengthening photoperiod promoting leaf senescence
  Rf[Rf > 0] = 0
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when photoperiod is shorter than the photoperiod threshold (P_base)
  # after the date of the longest photoperiod (summer solstice), namely, the 173rd day of year
  t0 <- vector()
  for(c in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Li[,c] < P_base)]
    ind1 = min(which(t0A > 173))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  
  # nullify values before the t0
  for(c in 1:ncol(data$Tmini)){
    Rf[1:t0[c],c] = 0 #nullify values before the date of leaf.out
  }
  
  # calculate the summation along the year (interval = 1:366) and derive the date of leaf.off
  # DOY of budburst criterium
  doy = apply(Rf,2, function(xt){
    doy = which(cumsum(xt) <= F_crit)[1]
  })
  
  return(doy)
}
DM2.model = function(par, data){
  # exit the routine as some parameters are missing
  if (length(par) != 3){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the
  # par argument for readability
  T_base = par[1]
  P_base = par[2]
  F_crit = par[3]
  
  # create forcing/chilling rate vector at the day level
  Rf = (data$Tmini - T_base)*(1-(data$Li/P_base)) # shortening photoperiod promoting leaf senescence
  Rf[Rf > 0] = 0
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when photoperiod is shorter than the photoperiod threshold (P_base)
  # after the date of the longest photoperiod (summer solstice), namely, the 173rd day of year
  t0 <- vector()
  for(c in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Li[,c] < P_base)]
    ind1 = min(which(t0A > 173))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  
  # nullify values before the t0
  for(c in 1:ncol(data$Tmini)){
    Rf[1:t0[c],c] = 0 #nullify values before the date of leaf.out
  }
  
  # calculate the summation along the year (interval = 1:366) and derive the date of leaf.off
  # DOY of budburst criterium
  doy = apply(Rf,2, function(xt){
    doy = which(cumsum(xt) <= F_crit)[1]
  })
  
  return(doy)
}
TPM.model = function(par, data){
  # exit the routine as some parameters are missing
  if (length(par) != 4){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the par argument for readability
  P_base = par[1]
  a = par[2]
  b = par[3]
  F_crit = par[4]
  
  # create forcing/chilling rate vector at the day level
  Rf = 1/(1+exp(a*(data$Tmini*data$Li-b)))
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when photoperiod is shorter than the photoperiod threshold (P_base)
  # after the date of the longest photoperiod (summer solstice), namely, the 173rd day of year
  t0 <- vector()
  for(c in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Li[,c] < P_base)]
    ind1 = min(which(t0A > 173))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  
  # nullify values before the t0
  for(c in 1:ncol(data$Tmini)){
    Rf[1:t0[c],c] = 0 #nullify values before the date of leaf.out
  }
  
  # calculate the summation along the year and derive the date of leaf.off
  # DOY of budburst criterium
  doy = apply(Rf,2, function(xt){
    doy = which(cumsum(xt) >= F_crit)[1]
  })
  
  return(doy)
}
SecondGen_PIA.models = function(par, predictor, data) {
  # exit the routine as some parameters are missing
  if (length(par) != 5 & length(par) != 6){
    stop("model parameter(s) out of range (too many, too few)")
  }
  
  # extract the parameter values from the
  # par argument for readability
  P_base = as.numeric(par[1])
  a = as.numeric(par[2])
  b = as.numeric(par[3])
  c = as.numeric(par[4])
  if(length(par)==5) {
    d = as.numeric(par[5])
    pred = predictor
  }
  if(length(par)==6) {
    d = as.numeric(par[5])
    e = as.numeric(par[6])
    pred1 = predictor[1]
    pred2 = predictor[2]
  }
  
  # create forcing/chilling rate vector at the day level
  Rf = 1/(1+exp(a*(data$Tmini*data$Li-b)))
  
  # photoperiod-dependent start-date for chilling accumulation (t0)
  # t0 is defined as the first day when photoperiod is shorter than the photoperiod threshold (P_base)
  # after the date of the longest photoperiod (summer solstice), namely, the 173rd day of year
  t0 <- vector()
  for(col in 1:ncol(data$Tmini)) {
    interval = 1:366
    t0A = interval[which(data$Li[,col] < 173)]
    ind1 = min(which(t0A > 173))
    t0A = t0A[ind1]
    t0 = c(t0,t0A)
  }
  
  # nullify values before the t0
  for(col in 1:ncol(data$Tmini)){
    Rf[1:t0[col],col] = 0 
  }
  
  if(length(par)==5) {
    # add predictor at the end of the matrix-columns
    Rf = rbind(Rf,predictor)
    
    # predict date of leaf.off
    doy = apply(Rf,2, function(xt){
      doy = which(cumsum(xt[1:366]) >= c+d*xt[367])[1]
    }) 
  }
  if(length(par)==6) {
    # add predictors at the end of the matrix-columns
    Rf = rbind(Rf,pred1)
    Rf = rbind(Rf,pred2)
    
    # predict date of leaf.off
    doy = apply(Rf,2, function(xt){
      doy = which(cumsum(xt[1:366]) >= c+d*xt[367]+e*xt[368])[1]
    }) 
  }
  
  return(doy)
}
df_sampled
```

## Forcing data: Min temperature and Photoperiod
This forcing data are obtained from cru (min temp) using the ingestr package and the geosphere pkg (photoperiod)
```{r}
## Calculate Tmin from CRU with ingestr a site ensemble
siteinfo <- unique(df_sampled[,c(3,4,5)]) %>% rename(sitename=id_site) %>% mutate(sitename=as.character(sitename)) %>%
  mutate(year_start = 1949, year_end = 2015,date_start = lubridate::ymd(paste0(year_start, "-01-01")),
         date_end = lubridate::ymd(paste0(year_end, "-01-02"))) %>% mutate(year_start=as.character(year_start)) %>% 
  mutate(year_end=as.character(year_end))

siteinfo <- as_tibble(siteinfo)

tmin_sample_cru <- ingest(
  siteinfo = siteinfo,
  source    = "cru",
  getvars   = list(temp = "tmin"),
  dir       = "~/data/cru/ts_4.01/"  # adjust this with your local path
)
save(tmin_sample_cru, file = "~/data/tmin_sample_cru.RData")

load("~/data/tmin_sample_cru.RData")
tmin_sample_cru$data[[1]]

tmin <- tmin_sample_cru %>% unnest(cols=c(data))
#tmin <- tmin %>% pivot_wider(names_from = sitename, values_from = tmin)
#tmin <- data.matrix(tmin)
tmin <- tmin %>% mutate(doy=paste0(yday(date),"tmin"),year=year(date)) %>% rename(id_site=sitename) %>% mutate(id_site=as.integer(id_site)) %>% dplyr::select(-date)
tmin <- tmin %>% pivot_wider(names_from = doy, values_from = tmin) 
tmin.df <- tmin %>% inner_join(df_sampled[,c(3,4,5,7)]) %>% distinct() %>% relocate(year,id_site,lon,lat) %>% arrange(id_site,year)

## Calculate photoperiod
#dl <- daylength(45, 1:365)
#tapply(dl, rep(1:12, c(31,28,31,30,31,30,31,31,30,31,30,31)), mean)

all.latitudes <- unique(df_sampled$lat)
df_photo <- list()
for(i in 1:length(all.latitudes)) {
  dl <- daylength(all.latitudes[i], 1:366)
  df_photo[[i]] <- dl 
}
length(unique(df_sampled$year))
length(unique(df_sampled$lat)) # include ncol
length(unique(df_sampled$id_site))

photo <- matrix(unlist(df_photo), ncol = length(unique(df_sampled$lat)))
photo <- as.data.frame(photo)
colnames(photo) <- all.latitudes
photo$doy <- paste0(1:366,"photo")
photo <- photo %>% pivot_longer(!doy, names_to = "lat", values_to = "photoperiod") %>% arrange(lat) %>% mutate(lat=as.double(lat))
photo <- photo %>% pivot_wider(names_from = doy, values_from = photoperiod) 
photo.df <- photo %>% full_join(df_sampled[,c(3,4,5,7)]) %>% distinct() %>% relocate(year,id_site,lon,lat) %>% arrange(id_site,year)

forcing <- tmin.df %>% inner_join(photo.df)

```

## Import & Format data
```{r}
# Phenological observations
pheno.df <- df_sampled %>% dplyr::select("id_species_site","id_site","lon","lat","species","year","off","on")

# Predictors
preds.df <- df_sampled %>% dplyr::select("id_species_site","id_site","lon","lat","species","year","on","temp_GS","RD_summer","cGSI","cA_tot","cA_tot-w")

# Define model names
models   <- c("CDD","DM1","DM2","TPM","SIAM","TDM","TPDM","PIA_gsi","PIA-","PIA+")

# Define species
sps <- unique(pheno.df$species)

# Define timeseries
id_species_site <- unique(pheno.df$id_species_site)

# Prepare input datasets for PHENOR package
phenor_input <- pheno.df %>% inner_join(forcing)

# Create a DataList to store all subsets for each sps
DataList <- replicate(length(sps), data.frame())
names(DataList) <- paste0("DataList","_",sps)
SiteList <- replicate(length(sps), data.frame())

# Initiate an external loop to subset for each sps
for(i in 1:length(sps)) {
  
  # Subset phenological dataset for each sps
  pheno_sp.sub <- phenor_input[which(phenor_input$species==sps[i]),]
  pheno_sp.sub <- as.data.frame(pheno_sp.sub)

  # Find the sites per sps
  sites <- unique(pheno_sp.sub$id_site)
  SiteList[[i]] <- sites
  
  # Add empty subsets
  DataList[[i]] <- replicate(length(sites),data.frame())
  names(DataList[[i]]) <- paste0("data_",sites,"_",pheno_sp.sub[1,]$species)
  
  # Counter for sites
  count  <- 0 
  
  # Loop for each site
  for(site in sites){
    index <- which(pheno_sp.sub$id_site==site)
    data = list("doy" = as.vector(pheno_sp.sub[index,]$on),
                "site" = as.vector(paste0(pheno_sp.sub[index,]$id_site,"_",pheno_sp.sub[1,]$species)),
                "location" = t(pheno_sp.sub[index,c("lon","lat")]), 
                "year" = as.vector(pheno_sp.sub[index,]$year),
                "Ti" = NULL,
                "Tmini" = t(pheno_sp.sub[index,paste0(1:366,"tmin")]),
                "Tmaxi" = NULL,
                "Li" = t(pheno_sp.sub[index,paste0(1:366,"photo")]),
                "SPEI" = NULL,
                "VPDi" = NULL,
                "transition_dates" = as.vector(pheno_sp.sub[index,]$off),
                "georeferencing" = NULL
    )
    
    # Store each site-specific dataframe in the DataList of the corresponding sps
    count <- count+1
    DataList[[i]][[count]] <- data
  }
  print(paste0(sps[i]," DONE!"))
}
```

```{r}
##----------------------------------------
## Predictions

data.sub <- DataList[[i]][[count]]

# Initialize dataframe
DoYoff_Preds.df <- data.frame() 
opt_pars.df <- data.frame() 

for(i in 1:2) { #length(species)
  
  # Initialize datasets per species
  DoYoff_Preds.sp <- data.frame()
  opt_pars.sp <- data.frame()
  
  # Calculate number of sites per species
  sites <- SiteList[[i]]
  
  for(site in 1:2) { #length(sites)
    
    # Subset according to id_species_site
    ts <- paste0(sps[i],"_",sites[site]) #paste0(sites[site],"_",sps[i])
    data.sub <- DataList[[i]][[site]]
    preds.sub <- preds.df %>% dplyr::filter(id_species_site==ts)
    
    # Initialize sub-dataframes to store results
    DoYoff_Preds.sub <- data.frame(id_species_site=data.sub$site, species=sps[i], year=data.sub$year)
    DoYoff_Preds.sub$Obs_DoYoff <- data.sub$transition_dates
    opt_pars.sub <- data.frame(id_species_site=data.sub$site, species=sps[i], id_site=site) #sites
    
    ## Parameter optimization and Prediction of leaf senescence dates
    # PHENOR package (Hufkenset al., 2018)

    
    ## CDD model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "CDD.model",
                                        method = "GenSA",
                                        lower = c(15,-3000),
                                        upper = c(30,0),
                                        control = list(max.call = 40000))
    opt_pars.sub$Tbase_CDD <- optimal_pars[1]
    opt_pars.sub$Fcrit_CDD <- optimal_pars[2]
    DoYoff_Preds.sub$Pred_DoYoff_CDD <- pr_predict(par = optimal_pars$par,
                                                           data = data.sub,
                                                           model = "CDD.model")
    
    ## DM1 model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "DM1.model",
                                        method = "GenSA",
                                        lower = c(15,11,-2000),
                                        upper = c(30,16,0),
                                        control = list(max.call = 40000))
    opt_pars.sub$Tbase_DM1 <- optimal_pars[1]
    opt_pars.sub$Pbase_DM1 <- optimal_pars[2]
    opt_pars.sub$Fcrit_DM1 <- optimal_pars[3]
    DoYoff_Preds.sub$Pred_DoYoff_DM1 <- pr_predict(par = optimal_pars$par,
                                                           data = data.sub,
                                                           model = "DM1.model")
    
    ## DM2 model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "DM2.model",
                                        method = "GenSA",
                                        lower = c(15,11,-2000),
                                        upper = c(30,16,0),
                                        control = list(max.call = 40000))
    opt_pars.sub$Tbase_DM2 <- optimal_pars[1]
    opt_pars.sub$Pbase_DM2 <- optimal_pars[2]
    opt_pars.sub$Fcrit_DM2 <- optimal_pars[3]
    DoYoff_Preds.sub$Pred_DoYoff_DM2 <- pr_predict(par = optimal_pars$par,
                                                           data = data.sub,
                                                           model = "DM2.model")
    
    ## TPM model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "TPM.model",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0),
                                        upper = c(16,0.1,250,200),
                                        control = list(max.call = 40000))
    opt_pars.sub$Pbase_TPM <- optimal_pars[1]
    opt_pars.sub$a_TPM <- optimal_pars[2]
    opt_pars.sub$b_TPM <- optimal_pars[3]
    opt_pars.sub$Fcrit_TPM <- optimal_pars[4]
    DoYoff_Preds.sub$Pred_DoYoff_TPM <- pr_predict(par = optimal_pars$par,
                                                           data = data.sub,
                                                           model = "TPM.model")
    
    ## SIAM model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        predictor = preds.sub$on,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "SecondGen_PIA.models",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0,0),
                                        upper = c(16,0.1,250,300,1),
                                        control = list(max.call = 40000))
    opt_pars.sub$Pbase_SIAM <- optimal_pars[1]
    opt_pars.sub$a_SIAM <- optimal_pars[2]
    opt_pars.sub$b_SIAM <- optimal_pars[3]
    opt_pars.sub$c_SIAM <- optimal_pars[4]
    opt_pars.sub$d_SIAM <- optimal_pars[5]
    DoYoff_Preds.sub$Pred_DoYoff_SIAM <- pr_predict(par = optimal_pars$par,
                                                            predictor = preds.sub$on,
                                                            data = data.sub,
                                                            model = "SecondGen_PIA.models")
    
    ## TDM model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        predictor = preds.sub$temp_GS,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "SecondGen_PIA.models",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0,0),
                                        upper = c(16,0.1,250,300,1),
                                        control = list(max.call = 40000))
    opt_pars.sub$Pbase_TDM <- optimal_pars[1]
    opt_pars.sub$a_TDM <- optimal_pars[2]
    opt_pars.sub$b_TDM <- optimal_pars[3]
    opt_pars.sub$c_TDM <- optimal_pars[4]
    opt_pars.sub$d_TDM <- optimal_pars[5]
    DoYoff_Preds.sub$Pred_DoYoff_TDM <- pr_predict(par = optimal_pars$par,
                                                           predictor = preds.sub$temp_GS,
                                                           data = data.sub,
                                                           model = "SecondGen_PIA.models")
    
    ## TPDM model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        predictor = c(preds.sub$temp_GS,preds.sub$RD_summer),
                                        data = data.sub,
                                        cost = rmse,
                                        model = "SecondGen_PIA.models",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0,0,0),
                                        upper = c(16,0.1,250,300,1,1),
                                        control = list(max.call = 40000))
    opt_pars.sub$Pbase_TPDM <- optimal_pars[1]
    opt_pars.sub$a_TPDM <- optimal_pars[2]
    opt_pars.sub$b_TPDM <- optimal_pars[3]
    opt_pars.sub$c_TDPM <- optimal_pars[4]
    opt_pars.sub$d_TPDM <- optimal_pars[5]
    opt_pars.sub$e_TPDM <- optimal_pars[6]
    DoYoff_Preds.sub$Pred_DoYoff_TPDM <- pr_predict(par = optimal_pars$par,
                                                            predictor = c(preds.sub$temp_GS,preds.sub$RD_summer),
                                                            data = data.sub,
                                                            model = "SecondGen_PIA.models")
    
    ## PIA_gsi model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        predictor = preds.sub$cGSI,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "SecondGen_PIA.models",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0,0),
                                        upper = c(16,0.1,250,300,1),
                                        control = list(max.call = 40000))
    opt_pars.sub$Pbase_PIAgsi <- optimal_pars[1]
    opt_pars.sub$a_PIAgsi <- optimal_pars[2]
    opt_pars.sub$b_PIAgsi <- optimal_pars[3]
    opt_pars.sub$c_PIAgsi <- optimal_pars[4]
    opt_pars.sub$d_PIAgsi <- optimal_pars[5]
    DoYoff_Preds.sub$Pred_DoYoff_PIAgsi <- pr_predict(par = optimal_pars$par,
                                                              predictor = preds.sub$cGSI,
                                                              data = data.sub,
                                                              model = "SecondGen_PIA.models")
    
    ## PIA- model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        predictor = preds.sub$`cA_tot-w`,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "SecondGen_PIA.models",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0,0),
                                        upper = c(16,0.1,250,300,1),
                                        control = list(max.call = 40000))
    opt_pars.sub$`Pbase_PIA-` <- optimal_pars[1]
    opt_pars.sub$`a_PIA-` <- optimal_pars[2]
    opt_pars.sub$`b_PIA-` <- optimal_pars[3]
    opt_pars.sub$`c_PIA-` <- optimal_pars[4]
    opt_pars.sub$`d_PIA-` <- optimal_pars[5]
    DoYoff_Preds.sub$`Pred_DoYoff_PIA-` <- pr_predict(par = optimal_pars$par,
                                                              predictor = preds.sub$`cA_tot`,
                                                              data = data.sub,
                                                              model = "SecondGen_PIA.models")
    
    ## PIA+ model
    optimal_pars <- pr_fit_parameters(par = NULL,
                                        predictor = preds.sub$cA_tot,
                                        data = data.sub,
                                        cost = rmse,
                                        model = "SecondGen_PIA.models",
                                        method = "GenSA",
                                        lower = c(11,0.02,100,0,0),
                                        upper = c(16,0.1,250,300,1),
                                        control = list(max.call = 40000))
    opt_pars.sub$`Pbase_PIA+` <- optimal_pars[1]
    opt_pars.sub$`a_PIA+` <- optimal_pars[2]
    opt_pars.sub$`b_PIA+` <- optimal_pars[3]
    opt_pars.sub$`c_PIA+` <- optimal_pars[4]
    opt_pars.sub$`d_PIA+` <- optimal_pars[5]
    DoYoff_Preds.sub$`Pred_DoYoff_PIA+` <- pr_predict(par = optimal_pars$par,
                                                              predictor = preds.sub$cA_tot,
                                                              data = data.sub,
                                                              model = "SecondGen_PIA.models")
    
    # Autumn anomalies
    DoYoff_Preds.sub$meansite_DoYoff <- mean(data.sub$transition_dates)
    DoYoff_Preds.sub$Obs_AnomDoYoff <- data.sub$transition_dates - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_CDD <- DoYoff_Preds.sub$Pred_DoYoff_CDD - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_DM1 <- DoYoff_Preds.sub$Pred_DoYoff_DM1 - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_DM2 <- DoYoff_Preds.sub$Pred_DoYoff_DM2 - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_TPM <- DoYoff_Preds.sub$Pred_DoYoff_TPM - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_SIAM <- DoYoff_Preds.sub$Pred_DoYoff_SIAM - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_TDM <- DoYoff_Preds.sub$Pred_DoYoff_TDM - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_TPDM <- DoYoff_Preds.sub$Pred_DoYoff_TPDM - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$Preds_AnomDoYoff_PIAgsi <- DoYoff_Preds.sub$Pred_DoYoff_PIAgsi - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$`Preds_AnomDoYoff_PIA-` <- DoYoff_Preds.sub$`Pred_DoYoff_PIA-` - DoYoff_Preds.sub$meansite_DoYoff
    DoYoff_Preds.sub$`Preds_AnomDoYoff_PIA+` <- DoYoff_Preds.sub$`Pred_DoYoff_PIA+` - DoYoff_Preds.sub$meansite_DoYoff
    
    # Bind site-datasets
    DoYoff_Preds.sp <- rbind(DoYoff_Preds.sp,DoYoff_Preds.sub)
    opt_pars.sp <- rbind(opt_pars.sp,opt_pars.sub)
    print(paste0("RUNNING: ",DoYoff_Preds.sub$id_species_site," ",site," OF ",length(sites)))
  }
  # Bind final datasets
  DoYoff_Preds.df <- rbind(DoYoff_Preds.df,DoYoff_Preds.sp)
  opt_pars.df <- rbind(opt_pars.df,opt_pars.sp)
}
write.table(DoYoff_Preds.df,"ModelAnalysis_1_Predicted_DoYoff.csv",sep=";",row.names = F)
write.table(opt_pars.df,"ModelAnalysis_2_OptimalParameters.csv",sep=";",row.names = F)

```


```{r}
##----------------------------------------
## Model Performance

## FIGURE 3A
# Observed (Obs_AnomDoYoff) vs.
# Predicted (Pred_AnomDoYoff)
# leaf senescence anomalies, i.e., as deviation from the mean observed leaf-out date at each site
# of the best-performing first-generation (TPM) [Lang et al. (2019)], 
# second-generation (TPDM) [Liu et al. (2019)]
# and CarbLim models (PIAM)

# Import data
DoYoff_Preds.df <- fread("ModelAnalysis_1_Predicted_DoYoff.csv")
DoYoff_Preds.df <- DoYoff_Preds.df %>% 
  filter(species != "Betula pubescens")

# Select autumn anomalies from best-performing models
best_pred_DoYoff <- DoYoff_Preds.df %>% 
  select(Obs_AnomDoYoff,Pred_AnomDoYoff_TPM,Pred_AnomDoYoff_TPDM,`Pred_AnomDoYoff_PIA+`)

# Format dataset for plotting
types <- c("First-generation (TPM)","Second-generation (TPDM)","CarbLim model (PIA+)")
colnames(best_pred_DoYoff) <- c("observations",types)
best_pred_DoYoff <- best_pred_DoYoff %>% 
  pivot_longer(-observations,names_to="model_type",values_to="predictions")
best_pred_DoYoff$model_type <- factor(best_pred_DoYoff$model_type, levels=types)

# Define palette
paletteBlueRed <- c("blue3","white","red3")

# Plot
fig_3a <- ggplot(best_pred_DoYoff, aes(x=predictions, y=observations)) +
  stat_bin2d(bins=235) +
  labs(x = "Predicted autumn anomaly",
       y = "Observed autumn anomaly") +
  coord_cartesian(xlim=c(-50,50), ylim=c(-50,50))+
  scale_fill_gradientn(colours = paletteBlueRed,
                       limits = c(10,2300),
                       breaks = c(100,1200,2200)) +
  geom_abline(slope=1, intercept=0,
              na.rm = FALSE, show.legend = NA, linetype="dashed") +
  theme(aspect.ratio=1,
        legend.position = "right",
        plot.title=element_text(hjust=.5),
        axis.title.y=element_text(size=14),
        axis.text.y=element_text(size=11),
        axis.title.x=element_text(size=14),
        axis.text.x=element_text(size=11),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = 'white', colour = 'black'),
        plot.margin = unit(c(0, 0, 0, 0), "cm")
  ) + 
  facet_wrap(.~model_type, ncol=3)
fig_3a

# Add R^2 values
# Average across time-series
# See Figure 2B
dat_text <- data.frame(
  label = c("R2 = 0.21", "R2 = 0.62", "R2 = 0.78"), #check R2_models (Figure 3B)
  model_type = types
)
fig_3a  <- fig_3a + geom_text(
  data = dat_text,
  mapping = aes(x=-35, y=45, label=label)
)

# Calculate intercept plus slope by Standard Major Axis (SMA)
SMA_values <- best_pred_DoYoff %>%  
  group_by(model_type) %>% 
  summarise(int=lmodel2(observations ~ predictions)$regression.results$Intercept[3],
            slope=lmodel2(observations ~ predictions)$regression.results$Slope[3])
fig_3a <- fig_3a +
  geom_abline(data = SMA_values,
              mapping = aes(intercept=int,slope=slope),
              linetype = "solid")
fig_3a

```

